
\documentclass[a4paper,12pt]{article}
\usepackage[T1]{fontenc}
\usepackage[latin1]{inputenc}
\usepackage{lmodern}

\title{The \textsf{fjodor} package\protect\footnote{This package 
is dedicated of the glory work of russian writers Fjodor Mihajlovi\v{c} Dostojevski.}}
\author{Zoran T. Filipovi\'{c} \\ 
        Jurija Gagarina 263/6 \\ 
        11070 Novi Beograd, Srbija}

\usepackage[english,serbian]{babel}
\def\dateserbian{%
  \def\today{\number\day .~\ifcase\month\or
    januar\or februar\or mart\or april\or maj\or
    jun\or jul\or avgust\or septembar\or oktobar\or 
    novembar\or decembar\fi \space \number\year.}}       

\begin{document}
\frenchspacing
\maketitle
\selectlanguage{english}
\begin{abstract}
This package produce a very nice layout and typeout for a small books. Working 
under a \verb|memoir| documentclass. Package contains several options and name 
of options is same name of writers of this books. 
\end{abstract}

\section{Put it!}

If you want to yous this package just put it the following line in your preambula:
\verb|\usepackage[options]{fjodor}| and push your machine.

\subsection{Options}

This package contains several options which is describle this.

\begin{description}
\item[fjodor] which i dadicated in Fjodor Mihailovi\v{c} Dostojevski produce layout in \verb|\pagemlargecrownvo| typeout in 9.5$\times$15.5\,cm whith 1.5\,cm spine margins 
and 2.5\,cm upper margins. Usefull for 10pt font size.
\item[katajev] which i dadicated in Pavel Valentinovi\v{c} Katajev produce layout in
17.5$\times$11.7\,cm, typeout in 9.1$\times$14.3\,cm whidth 1.3\,cm spine and upper margin. 
Total number of text lines is 34.
\item[hese] which i dadicated to Hermann Hesse produce layout in 19.8$\times$14\,cm
typeout in 9.8$\times$14.3\,cm whidth 1.75\,cm spine margin and 2.1\,cm upper margin.
Total number of text lines is 34.
\item[puskin] which i dadicated to Pu\v{s}kin produce layout in 19.8$\times$14\,cm
typeout in 9.9$\times$15.2\,cm whidth 2\,cm spine margin and 2.5\,cm uper margin.
\item[ernest] which i dadicated to Ernest Hemingway produce layout in 19$\times$12\,cm
typeout in 9$\times$14.5\,cm whidth 1.3\,cm spine margin and 1.7\,cm upper margin.
Total number of text lines is 35.
\item[viktor] which i dadicated to Viktor Jorofejev produce layout in 19$\times$12\,cm
typeout in 9.4$\times$15.04\,cm whidth 1\,cm spine margin and 1.4\,cm upper margin.
Total number of text lines is 36.
\item[remark] which i dadicated to Erich Maria Remarque produce layout in 19.5$\times$13.5\,cm
typeout in 9.9$\times$15.4\,cm whidth 1.3\,cm spine margin and 1.5\,cm upper margin.
Total number of text lines is 37.
\item[tehnicar] which i dadicated to small technical book, which was printing on Mechanical Faculity of Belgrade, produce layout in 20$\times$14\,cm
typeout in 10$\times$16\,cm whidth 1.5\,cm spine margin and 1.7\,cm upper margin.
Total number of text lines is 46. Usefull for small tehnical book in 9pt font size.
\item[ulro] which i dadicated in \v{C}eslav Milo\v{s}, book ZEMLJA ULRO, 
produce layout in 20$\times$14\,cm typeout in 9.8$\times$15.2\,cm whidth 2\,cm spine 
margin and 2\,cm upper margin. Total number of text lines is 37. Usefull for book in 10pt 
font size.
\item[aristotel] which i dadicated to greek philosophist Aristotel, book FIZIKA, 
produce layout in 20.8$\times$13.7\,cm typeout in 8.1$\times$16.7\,cm whidth 2.7\,cm spine 
margin and 2\,cm upper margin. Total number of text lines is 40. Usefull for book in 10pt 
font size.
\item[jaeger] which i dadicated to Charles Jaeger, book Th\'{e}orie G\'{e}n\'{e}rale du Coup
De B\'{e}lier, produce layout in 24.1$\times$15.6\,cm typeout in 10.4$\times$16.9\,cm 
whidth 2.3\,cm spine margin and 3\,cm upper margin. Total number of text lines is 41 for 10pt 
font size.
\item[tesla] which i dadicated to Nikola Tesla, book Colorado Springs Notes 1899---1900, 
produce layout in 27.7$\times$20.5\,cm typeout in 13.5$\times$21\,cm whidth 2.6\,cm spine 
margin and 2.6\,cm upper margin. Total number of text lines is 41 for 11pt font size.
\end{description}

\end{document}