
\documentclass[a4paper,12pt]{memoir}
\usepackage[T1]{fontenc}
\usepackage[latin1]{inputenc}
\usepackage{lmodern}
\usepackage{verbatim}
\usepackage{soul}
\usepackage{xspace}

\title{The \textsf{srbook-mem} package}
\author{Zoran T. Filipovi\'{c} \\ 
        Jurija Gagarina 263/6 \\ 
        11070 New Belgrade, Serbia}
        
\usepackage[serbian,english]{babel}
\usepackage[wilson]{srbook-mem}

%%% NICE WIDHTOUT WILSONDOB OPTIONS IN SRBOOK-MEM (NO LIGHT)
%%% \usepackage{srbook-mem}

%\chapterstyle{default}
%\chapterstyle{companion}
%\chapterstyle{chappell}
%\chapterstyle{ell}
%\chapterstyle{ger}
%\chapterstyle{ntglike}
%\chapterstyle{thatcher}

%%% NICE WIDTH WILSONDOB OPTIONS IN SRBOOK-MEM (NO LIGHT)
%%% \usepackage[wilsondob]{srbook-mem}

\chapterstyle{wilsondob}
%\chapterstyle{komalike} 
%\chapterstyle{crosshead}
%\chapterstyle{section}
%\chapterstyle{article}
%\chapterstyle{tandh}
%\chapterstyle{section}

\begin{document}
\frenchspacing
\maketitle

\begin{abstract}
In this document I try to explain how to produce a number of the book, part and chapter in
letters for serbian language in latin scripts. This package working very nice under \verb|memoir|
documentclass. 
\end{abstract}

\section*{Introduction}

If you defines serbian language (latin scripts) in babel package, and type 
\verb|\book{},\part{},\chapter{}| that produce, in  \verb|memoir| class, something like that: \textbf{Book I}, \textbf{Deo I} and \textbf{Glava I}. In many book in serbian language, and for many publisher we must produce something like that: 
\textbf{Knjiga Prva} like \textbf{First Book}, \textbf{Deo Prvi} like \textbf{First Part} 
and \textbf{Glava Prva} like \textbf{Chapter First}. 

In this meaning, I produce this package which working with \verb|memoir| documentclass.

\section*{The package}

If you want use this packages, you just put couple line like:

\begin{verbatim}
\documentclass[a4paper,12pt]{memoir}
\usepackage[T1]{fontenc}
\usepackage[latin1]{inputenc}
\usepackage[serbian]{babel}
\usepackage{srbook-mem}
\begin{document}
your text
\end{document}
\end{verbatim}

and that is produce \verb|\book| to the ten numbers (in letters), \verb|\part| to
the twenty numbers (in letters) and \verb|\chapter| to the fourty numbers (in letters).

Package provide a options which I call \verb|light| which produce Knjiga Prva, Deo Prvi,
Glava Prva, width \verb|normalfont|. The default font is bold. I find that is options 
\verb|light| is very nice and smooth. However, because is \verb|memoir| class is such
to many predefines chapterstyle, you may choose just a package, typing
\verb|\usepackage{srbook-mem}|, width no \verb|light| options. I find that the output is very nice for the following chapterstyle:
\begin{center}
\verb|default:companion:chappell:ell:ger:ntglike:thatcher|
\end{center}

That is produce number in letter in chapter for serbian language in latin script. 

The second options is to typing \verb|\usepackage[wilson]{srbook-mem}|, which produce 
Jedan, Dva, \dots{} like One, Two \dots{} I find that the output is very nice for the 
following chapterstyle:
\begin{center}
\verb|wilsondob:komalike:crosshead:section:article:tandh:section|
\end{center}

So, if you want to use this options you must 
typing, in preambula, line \verb|\usepackage[wilson]{srbook-mem}| push you machine 
to going, and the output is:

\selectlanguage{serbian}

\book{Ovo je prva knjiga}

\part{Ovo je prvi deo}
\selectlanguage{serbian}
\chapter{Ovo je prva glava}

However, you just see the \verb|Test.pdf| and \verb|TestLight.pdf| and---for 
my \TeX friend---evrything will be clear.

\chapter{Ovo je druga glava}

Or you just see this \verb|SerbianBookMem.tex| files, and \dots

\chapter{Ovo je tre\'{c}a glava}

Happy \TeX ing! In serbian language, of course!

\end{document}