\documentstyle[12pt,bg]{article}

\begin{document}

\setlength{\parindent}{0pt}
\setlength{\parskip}{1em}

This is a discussion of a position from the newsgroup rec.games.backgammon:

\begin{position}
\smallboard
\whitepoint{2}{2}
\whitepoint{3}{2}
\whitepoint{4}{2}
\whitepoint{5}{2}
\whitepoint{6}{5}
\whitepoint{9}{1}
\whitepoint{23}{2}
\blackpoint{1}{2}
\blackpoint{4}{2}
\blackpoint{5}{3}
\blackpoint{6}{2}
\blackpoint{8}{3}
\blackpoint{12}{1}
\blackpoint{24}{2}
\middlecube{1}

\boardcaption{Score 0-0 in a 3 point match: White to play 5-4}
\whiteonmove
\shownumbers
\showcube

\end{position}

{\bf Carl Nielsen (nlscar01@uctvax.uct.ac.za) wrote:}

I played 23-14. X rolled 4-2 and played 17-23. I couldn't get back on and 
ended up losing to a gammon. Did I play wrong? Should I have left moving my 
runners till later and moved 9-4 6-2 or something like that?

Suggestions please.

{\bf Kit Woolsey (kwoolsey@netcom.com) wrote:}

Your play looks ok to me.  You have to get moving sometime and your 
opponent has plenty of time if you play the waiting game, so it is 
probably best to run now before you have an accident and are forced to 
lose your board by rolling too many threes and fours.  You were just 
unlucky.  Btw, what is that cube doing in the center?  You have a clear 
advantage with some threats, so you have a strong double.  Maybe your 
opponent will pass, and maybe he should pass!

Kit

{\bf Kent Goulding (kg1@netcom.com) wrote:}

Biggest error here, by FAR, is the failure to double.  As is so often the 
case, the cube errors are the costly ones while checker plays are no big 
deal.  I think your opponent should pass.  As far as the checker play 
goes, it looks pretty close.  Running now gives your opponent a crack at 
grabbing TWO of your checkers, while waiting will likely expose only 
one.  Your home board is actually not in great danger of breaking right 
away. You can't play 5s and 6s there, and you will likely run from your 
anchor the next chance you get.  Also, it's nice to spread out some 
builders for your ace point in case your opponent runs (either forced or 
by choice).  I'm inclined to sit tight and see what happens next.

--KG

\end{document}
