%%% Testing:colours, borders and margins
\documentclass{article}

\usepackage{textpos}
\usepackage{color}

\setlength{\TPHorizModule}{20pt}
\setlength{\TPVertModule}\TPHorizModule

\setlength\parindent{0pt}
\pagestyle{empty}

\begin{document}

\definecolor{Blue}{rgb}{0.8,0.8,1.0}
\definecolor{Borders}{rgb}{0.8,1.0,0.8}

\begin{textblock}{4}(0,0)
  \textblockcolour{Blue}
  Blue
\end{textblock}

\begin{textblock}{12}(4,-1)
  \hrule width 12\TPHorizModule
  \vbox to 0pt{\hbox{12 units}\vss}
  \vskip1\TPVertModule
  \hrule width 8\TPHorizModule
  \vbox to 0pt{\hbox{8 units}\vss}
  \vskip1\TPVertModule
  \hrule width 4\TPHorizModule
  \vbox to 0pt{\hbox{4 units}\vss}
  \vskip1\TPVertModule
  \hrule width 2\TPHorizModule
  \vbox to 0pt{\hbox{2 units}\vss}
  \vskip1\TPVertModule
  \hrule width 1\TPHorizModule
\end{textblock}

\begin{textblock}{12}(4,4)
  This text is inside a box with a blue background.  The text is flush
  with the edge of the box, and the whole thing is twelve units wide.

  There's a second paragraph in here.
  \begin{itemize}
  \item Here is a list item.  It should be appropriately narrower than
  the text around it, and specifically not abut the border.
  \item Here is a second item.
  \end{itemize}
\end{textblock}

\TPshowboxestrue
\textblockrulecolour{red}
\setlength\TPboxrulesize{0.25\TPHorizModule}

\TPMargin{1\TPHorizModule}

\begin{textblock}{4}(0,7)
  \textblockcolour{Blue}
  Blue
\end{textblock}

\begin{textblock}{12}(4,11)
  This text is also inside a box with a blue background.  However,
  this time, there's a red border and a margin.  The box as a whole is
  twelve units wide, and the text is correspondingly narrower.
  \begin{itemize}
  \item Here is a list item.  It should be appropriately narrower than
  the text around it, and specifically not abut the border.
  \item Here is a second item.
  \end{itemize}
\end{textblock}

\TPMargin*{1\TPHorizModule}

\begin{textblock}{4}(0,19)
  \textblockcolour{Blue}
  Blue
\end{textblock}
\tekstblokroolkulur{Borders}

\begin{textblock}{12}(4,21)
  Finally, here is some text on a blue background with an
  unattractive green border.  This time, the text is twelve units wide,
  and the box, and its border, is consequently wider than that.
  \begin{itemize}
  \item Here is a list item.  It should be appropriately narrower than
  the text around it, and specifically not abut the border.
  \item Here is a second item.
  \end{itemize}
  Here's some final text, and a quotation:
  \begin{quotation}
  \noindent
  You shall see them on a beautiful quarto page, where a neat rivulet
  of text shall meander through a meadow of
  margin.
  \hfil\penalty50\hskip2em\hbox{}\nobreak\hfil
  \emph{--- Sheridan, The School for Scandal}\parfillskip=0pt
  \end{quotation}
\end{textblock}


\end{document}

