\documentclass[fontsize=13pt, DIV=8, pagesize=auto]{scrartcl}

\usepackage{fixltx2e}
\usepackage{etex}
\usepackage{xspace}
\usepackage{lmodern}
\usepackage[T1]{fontenc}
\usepackage{textcomp}
\usepackage{microtype}
\usepackage{hyperref}

\newcommand*{\mail}[1]{\href{mailto:#1}{\texttt{#1}}}
\newcommand*{\pkg}[1]{\textsf{#1}}
\newcommand*{\BibTeX}{Bib\kern-0.08em\TeX\@\xspace}

\pdfstringdefDisableCommands{%
  \def\BibTeX{BibTeX\xspace}%
}

\addtokomafont{title}{\rmfamily}

\title{The \pkg{lexikon} package\thanks{This manual corresponds to \pkg{lexikon}~v0.1c, dated~1999/05/13.}}
\author{Axel Kielhorn\\\mail{A.Kielhorn@tu-bs.de}}
\date{1999/05/13}


\begin{document}

\maketitle

\noindent
This is the \pkg{lexikon} package.
When I posted it to \texttt{de.comp.text.tex} some weeks ago it's name was 
\texttt{dictionary.sty}. I changed it to fit into 8~charakters.

This package provides some commands which may be useful for a two language 
dictionary with addiional information for the keywords. There is only 
\texttt{lexikon.tex} as a documentation and it is in german but should be sufficient 
to line out the usage. If you need phonetic symbols you have to get the 
\pkg{ips.sty} and read the manual, I can't help you with that.

This package is unsupported, but if you have an idea for an enhancement or 
find a bug just send me a mail. I'll try to fix it but won't spend many 
hours on the package because it already does what I need.

A database interface would be nice but will take some time.
My first idea was to use \BibTeX, but now i started using perl. I'm not sure 
It will ever be ready.

\end{document}


