\RequirePackage[!use=publish]{spdef}
\documentclass{article}
% var aSubmitFields = new Array("IdInfo", "Quiz1", "pathToTxtFile", "courseName", "quizName", "pdfFileName", "rtnURL");
% this.submitForm({cURL: "http://localhost/scripts/eq2db/eqText_html.asp", bEmpty: true, aFields: aSubmitFields, cSubmitAs: "HTML" });
% cPathToTxtFile = Request.Form("pathToTxtFile")
% cRtnURL = Request.Form("rtnURL")
% AcroTeX eDucation Bundle required for the email option.
%
\usepackage[online,coverpage,forcolorpaper,nopoints,nosolutions,useforms]{eqexam}
\usepackage[submitAs=html,eqText,tagged]{eq2db}

\title[TS]{\texorpdfstring{\TeX}{\textbackslash TeX} Usage Survey}
\author{D. P. Story}
\subject{\texorpdfstring{\sffamily eqExam}{eqExam} Demo}
\date{Spring, \the\year}
\keywords{TeX Usage Survey, eqExam, AcroTeX}

\university
{%
      THE UNIVERSITY OF AKRON\\
    Theoretical and Applied Mathematics
}
\email{dpstory@uakron.edu}

%\EmailSubject{TeX/LaTeX Survey Results}
%\EmailCourseName{eqExam Demo}
%\EmailExamName{TeX Usage Survey}
%\ServerRetnMsg{Your TeX survey results have been received, thank you.}
%\SubmitInfo{http://myWebSite/scripts/eqAttach.asp?nopath\#FDF}{myname@mymailprovider}

\ifpublish
\eqSubmit{http://faculty.nwfsc.edu/web/math/storyd/scripts/eqText_html.asp}%
    {d:/web/facultyweb/web/math/storyd/scripts/data/texsurvey.txt}{eqexam.TeXSurvey}
\else
\eqSubmit{http://localhost/scripts/eq2db/eqText_html.asp}%
    {c:/Inetpub/Data/texsurvey.txt}{eqexam.TeXSurvey}
\fi
\rtnURL{http://faculty.nwfsc.edu/web/math/storyd}

\lhead{\textsf{eqExam} Demo}
\rhead{\TeX{} Usage Survey}

\renewcommand\maketitledesign
{%
    \parbox{\textwidth}{\normalsize
        \makebox[\textwidth][c]{\bfseries\sffamily\Large\webtitle}\\[3ex]
        \makebox[\textwidth][s]{An \textsf{eqExam} Demo \hfill\webauthor}}
}
\makeatletter
\renewcommand\eqexcoverpagedesign
{%
    \thispagestyle{empty}
    \addtolength\textwidth{\oddsidemargin}
    \vspace*{.1\textheight}
    \noindent\hspace*{-\oddsidemargin}%
    \makebox[\linewidth]{\parbox{\linewidth}%
        {\bfseries\color{blue}\large\centering\webuniversity}}
    \par\vspace{.1\textheight}
    \noindent\hspace*{-\oddsidemargin}%
    \makebox[\linewidth]{\parbox{\linewidth}%
        {\bfseries\color{blue}\large\centering\websubject}}
    \par\vspace{\bigskipamount}
    \noindent\hspace*{-\oddsidemargin}%
    \makebox[\linewidth]{\parbox{\linewidth}%
        {\bfseries\color{blue}\large\centering\webtitle}}
    \par\vfill
    \optionalpagematter
    \par\vfill
    \parindent0pt
    \@date
}
\makeatother

%
% From the Web package, part of the AcroTeX eDucation Bundle
%
\newNaviIcon[j]{\GoToSubmit}{20pt}{15pt}{\textcolor{webblue}{$\blacktriangleleft$}}{this.pageNum=1}
\def\optionalpagematter
{%
    \begin{center}\setlength{\fboxrule}{1pt}
    \fcolorbox{blue}{webyellow}{%
    \begin{minipage}{.67\linewidth}
    \noindent\textcolor{red}{\textbf{Instructions:}} First, thank
    you very much for taking the time to fill out this survey of
    \TeX/\LaTeX{} usage. I will summarize the results, and make
    them publicly available. The survey has not been commissioned
    by TUG ({\TeX} User's Group), but is a demo of some of
    capabilities of my new {\LaTeX} package
    \textsf{eqExam}.\par\medskip When you have finished, click on
    the ``Submit'' button to record your responses.\par\medskip
    \textbf{\textcolor{red}{Note:}} The \textcolor{webblue}{Adobe
    Reader}, or an \textcolor{webblue}{Acrobat} application
    (version 5.0 or later) is required to submit this
    survey.\hfill\mbox{D. P. Story}
    \end{minipage}}
    \end{center}
}

\begin{document}

\maketitle

\begin{exam}{TeXSurvey}

\begin{instructions}[\TeX/\LaTeX.]
This first block of questions concerns your use of {\TeX/\LaTeX}.
\end{instructions}

\begin{problem*}
Experience using \TeX/\LaTeX.
\begin{parts}
\item How would you describe yourself in terms of experience using \TeX/\LaTeX?
\begin{answers}{4}
\rowsep{1bp}
\bChoices
    \Ans0 Novice\eAns
    \Ans0 Intermediate\eAns
    \Ans0 Experienced\eAns
    \Ans0 Package Developer\eAns
\eChoices
\end{answers}

\item I have used \TeX/\LaTeX{} for\dots
\begin{answers}{3}
\rowsep{1bp}
\bChoices
    \Ans0 less than a year\eAns
    \Ans0 one to two years\eAns
    \Ans0 two to five years\eAns
    \Ans0 five to ten years\eAns
    \Ans0 ten to fifteen years\eAns
    \Ans0 fifteen to twenty years\eAns
    \Ans0 more than twenty years\eAns
\eChoices
\end{answers}

\end{parts}
\end{problem*}

\begin{problem}
Approximately how many hours do you spend using \TeX/\LaTeX{} in a typical week?
On a per week basis, I use \TeX/\LaTeX\dots
\begin{answers}{4}
\rowsep{1bp}
\bChoices
    \Ans0 less than an hour\eAns
    \Ans0 one to two hours\eAns
    \Ans0 three to five hours\eAns
    \Ans0 ten hours\eAns
    \Ans0 more than ten hours\eAns
\eChoices
\end{answers}
\end{problem}

\begin{problem}
There are many {\TeX} user's groups worldwide, are you a member of one or more?
\begin{answers}{4}
\bChoices
    \Ans0 Yes\eAns
    \Ans0 No\eAns
\eChoices
\end{answers}
\end{problem}

\begin{problem*}
The following questions concern your preferred use of \TeX.
\begin{parts}
\item\label{whichTeX} Which styles to you predominately use?
\begin{answers}{4}
\rowsep{1bp}
\bChoices[2]
    \Ans0 Plain \TeX\eAns
    \Ans0 \LaTeX\eAns
    \Ans0 Con\TeX t\eAns
    \Ans0 Other, \makebox[0pt][l]{please specify \fillin{1.25in}{\empty}}\eAns
\eChoices
\end{answers}

\item Based on your response to \hyperref[whichTeX]{(\ref*{whichTeX})},
please explain why you made the choice you did.
\begin{solution}[.75in]
\end{solution}


\item There are many {\TeX} Systems available for use, commercial,
shareware, freeware. Which one do you primarily use?
\begin{answers}{4}
\rowsep{1bp}
\bChoices
    \Ans0 A \TeX{} Live System\eAns
    \Ans0 te\TeX\eAns
    \Ans0 Mik\TeX\eAns
    \Ans0 PC\TeX\eAns
    \Ans0 Y\negthinspace{\scriptsize\&}\negthinspace Y \TeX{} System\eAns
    \Ans0 Textures\eAns
    \Ans0 CMac\TeX\eAns
    \Ans0 True\TeX\eAns
    \Ans0 Oz\TeX\eAns
    \Ans0 Other, \makebox[0pt][l]{please specify \fillin{1.75in}{\empty}}\eAns
\eChoices
\end{answers}

\item What is your favorite {\TeX} editor? Specify \fillin{1.75in}{\empty}

\end{parts}
\end{problem*}

\begin{problem*}
What operating systems (OS) do you use \TeX?

\setlength\columnseprule{0pt}

\begin{parts}

\item What operating system (OS) runs on the machine you
most frequently use when working with \TeX/\LaTeX? Select one.
\begin{multicols}{2}
\begin{answers}{1}
\bChoices
     \Ans0 Windows XP\eAns
     \Ans0 Windows ME\eAns
     \Ans0 Windows NT\eAns
     \Ans0 Windows 2000\eAns
     \Ans0 Windows 98\eAns
     \Ans0 Windows 95\eAns
     \Ans0 Macintosh OS X\eAns
     \Ans0 Macintosh OS 9\eAns
     \Ans0 Macintosh OS 8\eAns
     \Ans0 Macintosh OS 7 or earlier\eAns
     \Ans0 Unix or Linux\eAns
     \Ans0 Other, please specify \fillin{1.25in}{\empty}\eAns
     \Ans0 Don't know\eAns
\eChoices
\end{answers}
\end{multicols}

\item Some people work on multiple platforms, what is your secondary
OS for working with \TeX/\LaTeX?
\begin{multicols}{2}
\begin{answers}{1}
\bChoices
     \Ans0 Windows XP\eAns
     \Ans0 Windows ME\eAns
     \Ans0 Windows NT\eAns
     \Ans0 Windows 2000\eAns
     \Ans0 Windows 98\eAns
     \Ans0 Windows 95\eAns
     \Ans0 Macintosh OS X\eAns
     \Ans0 Macintosh OS 9\eAns
     \Ans0 Macintosh OS 8\eAns
     \Ans0 Macintosh OS 7 or earlier\eAns
     \Ans0 Unix or Linux\eAns
     \Ans0 Other, please specify \fillin{1.25in}{\empty}\eAns
     \Ans0 Don't use two systems.\eAns
\eChoices
\end{answers}
\end{multicols}

\end{parts}

\end{problem*}


\begin{eqComments}[Seeking Help.]\end{eqComments}

\begin{problem*}
These questions concern any problems you may have with \TeX/\LaTeX{} and resources available to you.
\begin{parts}
\item  Do you struggle on alone (with \TeX/\LaTeX ) if things are going wrong, or do you
  seek help?
  \begin{answers}{3}
    \bChoices
        \Ans0 Struggle on along\eAns
        \Ans0 Seek help from a friend or colleague\eAns
        \Ans0 Seek help from a public forum\eAns
    \eChoices
  \end{answers}

\item Do you use on-line resources?
  \begin{answers}{4}
    \bChoices
        \Ans0 No\eAns
        \Ans0 Yes, I use \makebox[0pt][l]{\fillin{1.75in}{\empty}}\eAns
    \eChoices
  \end{answers}

\goodbreak

\item Do you seek help from \texttt{comp.text.tex}?
  \begin{answers}{3}
    \bChoices
        \Ans0 No\eAns
        \Ans0 Yes, I read the postings occasionally\eAns
        \Ans0 Yes, I read the postings regularly\eAns
        \Ans0 Yes, I even ask questions\eAns
        \Ans0 Yes, I answer questions on occasion\eAns
        \Ans0 Yes, I regularly answer questions\eAns
    \eChoices
  \end{answers}
\end{parts}
\end{problem*}

\begin{problem*}
The \textsf{CTAN} Archive.
\begin{parts}
\item\label{CTAN} Do you know about \textsf{CTAN} (Central {\TeX} Archive Network)?
  \begin{answers}{4}
    \bChoices
        \Ans0 No\eAns
        \Ans0 Yes\eAns
    \eChoices
  \end{answers}

\item If you answered ``Yes'' to question~\hyperref[CTAN]{(\ref*{CTAN})}, approximately, how often
      do you use this archive?
  \begin{answers}{3}
    \bChoices
        \Ans0 Once or twice a month\eAns
        \Ans0 three to five times per month\eAns
        \Ans0 More than five times per month\eAns
    \eChoices
  \end{answers}

\end{parts}
\end{problem*}

\begin{problem*}
The {\TeX} \textsf{FAQ} Resource.
\begin{parts}
\item\label{FAQ} Do you know there is a {\TeX} \textsf{FAQ} online resource? (\texttt{http://www.tex.ac.uk/faq})
  \begin{answers}{4}
    \bChoices
        \Ans0 No\eAns
        \Ans0 Yes\eAns
    \eChoices
  \end{answers}

\item\label{FAQ1} If you answered ``Yes'' to question, \hyperref[FAQ]{(\ref*{FAQ})} have you ever used the \textsf{FAQ}?
  \begin{answers}{4}
    \bChoices
        \Ans0 No\eAns
        \Ans0 Yes, \makebox[0pt][l]{how often? \fillin{1in}{\empty} per year.}\eAns
    \eChoices
  \end{answers}
\end{parts}
\end{problem*}

\begin{problem}
In terms of paper publications, what are the titles of the books you use for a reference to your
work in \TeX/\LaTeX? Please list titles/authors in the space provided below.
\begin{solution}[1.8in]
\end{solution}
\end{problem}

\goodbreak
\begin{eqComments}[Print and PDF.]\end{eqComments}

\begin{problem}
What application do you use to create PDF? Adobe Distiller, ps2pdf, other.
\begin{answers}{3}
\rowsep{1bp}
    \bChoices
        \Ans0 Adobe Distiller\eAns
        \Ans0 \textsf{ps2pdf} (\textsf{pdfwriter})\eAns
        \Ans0 \textsf{pdftex}\eAns
        \Ans0 \textsf{dvipdfm}\eAns
        \Ans0 Other, \makebox[0pt][l]{please specify \fillin{1.75in}{\empty}}\eAns
    \eChoices
\end{answers}
\end{problem}

\begin{problem}
Describe the reasons you create PDF document from a \TeX/\LaTeX{} source. Is it for
distribution over the Internet; to create interactive, high
content materials; for business purposes? Please elaborate in the
space provided below.
\begin{solution}[.75in]
\end{solution}
\end{problem}

\begin{problem}
Do you use {\TeX/\LaTeX} primarily for creating printable documents, or for creating
a PDF document suitable for publication on the Internet, or distribution to a colleague?
\begin{answers}{3}
    \bChoices
    \Ans0 Exclusively for print\eAns
    \Ans0 Mostly for PDF\eAns
    \Ans0 About half and half\eAns
    \eChoices
\end{answers}
\end{problem}

\begin{problem}
If you have any experience or knowledge of commercial publishing
houses that use a \TeX/\LaTeX{} or \textsf{pdftex} work flow  for
prepress typesetting,  please list publishing houses and countries
in the space provided below.
\begin{solution}[.75in]
\end{solution}
\end{problem}

\begin{eqComments}[Population Demographics.]
The next set of questions concern you.
\end{eqComments}

\begin{problem}
My gender is \dots
\begin{answers}{5}
    \bChoices
        \Ans0 Male\eAns
        \Ans0 Female\eAns
        \Ans0 \makebox[0pt][l]{Prefer not to disclose}\eAns
    \eChoices
\end{answers}
\end{problem}

\begin{problem}
I am \fillin{.75in}{\empty} years old.
\end{problem}

\begin{problem}
Which of the following best describes your status as
an \TeX{} user today? Select one:

\begin{answers}{1}
    \bChoices
        \Ans0 Use as a student\eAns
        \Ans0 Use as an educator/trainer\eAns
        \Ans0 Use at home for personal use or as a hobby\eAns
        \Ans0 Use for a home-based business\eAns
        \Ans0 Use in a company or business (non-home based)\eAns
    \eChoices
\end{answers}
\end{problem}

\begin{problem}
What is your country of residence? Select one:

\setlength\columnseprule{0pt}

\begin{multicols}{2}
\begin{answers}{1}
    \bChoices
        \Ans0 United States of America\eAns
        \Ans0 Canada\eAns
        \Ans0 Other Americas, \mbox{specify: \fillin{1.24in}{\empty}}\eAns
        \Ans0 United Kingdom\eAns
        \Ans0 France\eAns
        \Ans0 Germany\eAns
        \Ans0 Sweden\eAns
        \Ans0 Netherlands\eAns
        \Ans0 Other Europe, \mbox{specify: \fillin{1.25in}{\empty}}\eAns
        \Ans0 Japan\eAns
        \Ans0 Hong Kong/Taiwan\eAns
        \Ans0 China\eAns
        \Ans0 Other Asia, \mbox{specify: \fillin{1.25in}{\empty}}\eAns
        \Ans0 Australia\eAns
        \Ans0 New Zealand\eAns
        \Ans0 Africa, \mbox{specify: \fillin{1.25in}{\empty}}\eAns
    \eChoices
\end{answers}
\end{multicols}
\end{problem}

\def\myDing{\makebox[0pt][r]{\textcolor{red}{\ding{042}}\enspace}}

\begin{eqComments}[]%
\myDing If your company/organization uses \TeX/\LaTeX/\textsf{pdftex} in their business work flow,
please answer questions \textbf{\textcolor{blue}{\ref*{buss1}}} and \textbf{\textcolor{blue}{\ref*{buss2}}}.
\end{eqComments}

\begin{problem}\label{buss1}
How many employees work at your company including all
sites? Select one:

\setlength\columnseprule{0pt}
\begin{multicols}{3}
\begin{answers}{1}
    \bChoices
        \Ans0 One/work alone\eAns
        \Ans0 2 - 9\eAns
        \Ans0 10 - 49\eAns
        \Ans0 50 - 99\eAns
        \Ans0 100 - 499\eAns
        \Ans0 500 - 999\eAns
        \Ans0 1000 - 2999\eAns
        \Ans0 3000 - 4999\eAns
        \Ans0 5000 or more\eAns
        \Ans0 Don't know\eAns
    \eChoices
\end{answers}
\end{multicols}
\end{problem}

\begin{problem}\label{buss2}
What is your company/organization's PRIMARY industry? Select one:

\setlength\columnseprule{0pt}

\begin{multicols}{2}
\begin{answers}{1}
    \bChoices
        \Ans0 Aerospace\eAns
        \Ans0 Automotive\eAns
        \Ans0 Computer hardware\eAns
        \Ans0 Computer software\eAns
        \Ans0 Education\eAns
        \Ans0 Engineering\eAns
        \Ans0 Financial services\eAns
        \Ans0 Government\eAns
        \Ans0 Graphic design/fine art\eAns
        \Ans0 Healthcare/social services\eAns
        \Ans0 Insurance\eAns
        \Ans0 Legal\eAns
        \Ans0 Manufacturing\eAns
        \Ans0 Membership organizations - religious, nonprofit, trade unions\eAns
        \Ans0 Military\eAns
        \Ans0 Pharmaceuticals/bio tech\eAns
        \Ans0 Professional services (documentation)\eAns
        \Ans0 Professional services (technical/IT/Internet)\eAns
        \Ans0 Printing and publishing\eAns
        \Ans0 Public and protective services (police, fire)\eAns
        \Ans0 Retail\eAns
        \Ans0 Telecommunications\eAns
        \Ans0 Transportation\eAns
        \Ans0 Travel\eAns
        \Ans0 Wholesale\eAns
        \Ans0 Other, \mbox{please specify: \fillin{1.25in}{\empty}}\eAns
    \eChoices
\end{answers}
\end{multicols}
\end{problem}

\begin{eqComments}[Improvements in \TeX.]\end{eqComments}

\begin{problem}
What you most like to see improved in \TeX/\LaTeX? Use the space provided
below:
\begin{solution}[2in]
\end{solution}
\end{problem}


\begin{eqComments}[Thank You] once again for participating in this survey.\end{eqComments}
\begin{eqComments}[\GoToSubmit]
Click on the left arrow to return to the first page to submit your responses.\end{eqComments}



\end{exam}
\end{document}
