%\iffalse 
% local03.tex --- 01/09/93
%                 28/10/96
%
% Copyright (C) 1996 by Wenzel Matiaske, mati1831@perform.ww.tu-berlin.de
%
% As input to the local LaTeX-guide "`local.tex"'.
%
% For distribution of this document  see the copyright notice in the
% original sources mentioned below.
%
%\fi

\makeatletter
\newif\iflocalin
\newif\ifappendixin
\@ifundefined{localin}{\localintrue}{\localinfalse}
\@ifundefined{appendixin}{\appendixinfalse}{\appendixintrue}
\@ifundefined{docdir}{\def\docdir{\dots /emtex/doc/}}{}
\@ifundefined{bibtex}{\def\bibtex{%
{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}}{}
\@ifundefined{file}{\def\file#1{{\sl `#1'\/}}}{}
\makeatother

\iflocalin
  \subsubsection{The \bibtex\ Styles {\tt thesis}, {\tt journal}, and {\tt paper}}
\fi
\ifappendixin
  \subsection{The \bibtex\ Styles {\tt thesis}, {\tt journal}, and {\tt paper}}
\fi

The local style collection contains three `apalike' \bibtex-styles. The
first one is called \verb|thesis.bst| and produces a bibliography equal
to the format of the American Sociological Review. \verb|journal.bst|
equals the standard `apalike' format. The format \verb|paper.bst| is
designed for bibliographies in the style of the
``Handw\"{o}rterb\"{u}cher der Betriebs\-wirtschafts\-lehre'' (Gabler).
The \bibtex-styles are multilingual and support a {\tt language} field
in the bibliography database. 

These style files allow short citations via  the commands
\verb|\citeasnoun{|\emph{ref}\verb|}| or
\verb|\cite*{|\emph{ref}\verb|}| (e.~g. Author (year)),
\verb|\citeauthor|, and
\verb|\citeyear|.

The commands are defined in a corresponding \LaTeX{} package
called \verb|mlbib|. The package works together with {\tt
german.sty}, so the language can be changed with the
\verb|\selectlanguage{\|\emph{language}\verb|}| command. Even if
the {\tt language} field is used in the database, single lingual
bibliographies can be produced with the package option
\verb+singlelingual+.

The package also supports changes of the citation and bibliography
layout. The layout of the bibliography and the citations may be changed
with the package options \verb+thesis+, \verb+journal+, and
\verb+paper+ which provide some standard layout. These standard forms
may be changed via the commands 
\verb+\bblfont{}+,     
\verb+\bblauthorfont{}+,   
\verb+\bbltitlefont{}+,   
\verb+\bblvolumefont{}+,
\verb+\bblleft{}+, and   
\verb+\bblright{}+, which change the layout of the bibliography.
The layout of citations may
be modified with the commands
 \verb+\citeauthorfont{}+,   
 \verb+\citebetween{}+,           
 \verb+\citeleft{}+,                 
 \verb+\citeright{}+,                
 \verb+\citeasnounleft{}+ and       
 \verb+\citeasnounright{}+.         


\iflocalin
For more details see the german documentation
\file{\docdir thebib}.
\fi
