\chapter{Basic Use}

\section{Introduction}

In this section, I take the main file, show some snippets, and explain what
they all mean, or why you would want to use these things.

\section{example.tex}

\begin{verbatim}
\documentclass{uafthesis}
\end{verbatim}

This is where the secret sauce is.

\begin{verbatim}
\usepackage{fixltx2e} % Allows \(\) in captions, amongst other things.
\usepackage{ppl} % The Paladino font
\usepackage{amsmath, amssymb, amsfonts} % Thanks, AMS!
\usepackage{graphicx, float} % Graphics stuff
\usepackage{verbatim} % For very basic listings and multi-line comments.
%\usepackage{chapterbib} % This is an option for those bundling papers.
\usepackage[square]{natbib}
%\usepackage{tocbibind} % This fixes the "bibliography in ToC" problem.
                        % Use with chapterbib.
\usepackage{url} % I quote some URLs in the bibliography"
\end{verbatim}

\texttt{fixltx2e} fixes an annoying bug where using inline mathematics
delimiters in captions for graphics and tables would cause errors in the
compiler. Alternately, one may simply use the dollar signs instead.

\texttt{ppl} is the ``Paladino'' font, which has been extremely popular for
UAF theses in the past. There is, however, no rule against using ``Computer
Modern'' or some other font, and in fact ``Paladino'' was originally intended
for headings only when it was designed (fun fact).

Some writers of theses end up bundling multiple published papers into a thesis,
especially in the case of PhD candidates. \texttt{chapterbib} allows for
separate bibliographies for each chapter, which is the most appropriate format
given this bundled-paper style of thesis. Many theses---mine included---only
have a single bibliography.

Finally, \texttt{natbib} allows one to change how citations appear.

\small
\begin{verbatim}
\input{custom-macros.tex}
\end{verbatim}
\normalsize

Many authors write their own \LaTeX macros in order to make writing their thesis
easier. This document does not have any custom macros.

\small
\begin{verbatim}
\begin{document}

\title{An Example Thesis Document}
\author{Joshua Holbrook}

\degreeyear{2011}
\degreemonth{May}
\degree{Master of Arts}
\department{Dept. of Fresh Beats}
\numberofmembers{3} % Make sure this is right! The grad school hates empty
                    % signature lines.
\prevdegrees{B.A.M.F.}
\college{College of Pwning Noobs}

\makesig
\maketitle
\end{verbatim}
\normalsize

In this section, the important information about the thesis is filled in, and
then the signature page and title page are generated.

\small
\begin{verbatim}
% Wondering when to use `input' and when to use `include?'
% read http://en.wikibooks.org/wiki/LaTeX/Basics#Big_Projects .
\begin{abstract}
  \input abstract.tex
\end{abstract}

%Table of Contents and such
\tableofcontents
\listoffigures
\listoftables
%\listofothermaterials
\listofappendices
\end{verbatim}
\normalsize

Here, the abstract is inserted, then the table of contents and other tables.
Note that the appendices are in a separate table. \textbf{This is not typical}
in \LaTeX and requires special handling, as detailed in the next chapter.
Similarly with the List of Other Materials, if your thesis has one (think
CDs and such).

\small
\begin{verbatim}
\begin{acknowledgements}
  \input acknowledgements.tex
\end{acknowledgements}

\begin{quotepage}
  \input quotepage.tex
\end{quotepage}

\include{ch1}
\include{ch2}
\include{ch3}
\end{verbatim}
\normalsize

After inputting a few more pages, the chapters (separate documents) are all
included.

\small
\begin{verbatim}
\nocite{wikibook}
\bibliographystyle{agufull08}
\bibliography{thesis}
\end{verbatim}
\normalsize

This generates the bibliography. Note that the bibliography style is set to
\texttt{agufull08}. Generally, the graduate school isn't picky about 
bibliography style, as long as it's consistent. For geophysics papers (very
common at UAF), the AGU style is a great choice. It is included here, but may
also be found at AGU's web site.

\small
\begin{verbatim}
\appendix
\include{apx1}

\end{document}
\end{verbatim}
\normalsize

This is how appendices are included. In fact, they are written just like
regular chapters, and the \textbackslash appendix flag signals that following
chapters should be given letters (A, B, C...) instead of numbers (1, 2, 3...).
