\chapter{Gotchas and Caveats}

\section{Introduction}

\texttt{uafthesis} is not perfect. It still has issues that you need to keep
in mind.

\section{List of Appendices Chicanery}

In an ideal world, \texttt{uafthesis} would write your appendices to the
list of appendices instead of the table of contents. This is currently a bug,
and will be fixed in the future.

The workaround is this:

\begin{enumerate}
\item After generating a pretty-much-done thesis, open up the .toc file. In
this case, the file is called ``example.toc.'' Also open up a corresponding
.loa file (``example.loa'').
\item Cut the lines that refer to appendices out of the .toc file and paste
them into the .loa file. Save both. Make sure the first line in the .loa
that adds "Page" above the pages column stays put.
\item \emph{Make backups of both files.} This is because \LaTeX will overwrite
the .toc file in the next step.
\item Compile (\texttt{pdflatex example}) \emph{exactly once.}
\end{enumerate}

\section{List of Other Materials}

A similar process applies to the List of Other Materials as well:

\begin{quotation}
{ \it ``If you add a pocket or anything else to your thesis (like a CD-ROM)
then you have to have one more List in the Table of Contents.  Follow
the exact same procedure as in step 14, but now you must add the
command:

\textbackslash listofothermaterials

Notice that this goes before the \textbackslash listofappendices.  The file that you
must edit is the thesis12.lom file.  You have to create this by hand.
It is simple.  Here is the example:

\textbackslash renewcommand\textbackslash @pnumwidth\{3.55em\}
\textbackslash contentsline \{section\}{\textbackslash numberline \{A\}CD-ROM of Thesis, Defense and Sandpile Code\}\{Pocket\}
\textbackslash renewcommand\textbackslash @pnumwidth\{1.55em\}

Again, once you save this and run latex once, it will be erased.  A
good habit is to make thesis12.lom.bak and thesis12.loa.bak so they
will always be there.  Then copy them to thesis12.lom and thesis12.loa
and run latex your final time.

I did the \textbackslash @pnumwidth stuff above because the word `Pocket' is wider
than is allowed for a page number.  So I changed it just for this one
line and then changed it back to what it was (as found in the
uafthesis2004.cls file).''\\
\hspace*{1in}---Ryan Woodard
\end{quotation}

I suspect that there's a better way to do this, but I haven't gotten that far
yet.

\section{Bibliography Listings}

If you use \texttt{chapterbib}, also use \texttt{tocbibind} to make sure that
your bibliographies show up in the Table of Contents.
