\documentclass[DIV=9, pagesize=auto]{scrartcl}

\usepackage{fixltx2e}
\usepackage{etex}
\usepackage{lmodern}
\usepackage[T1]{fontenc}
\usepackage{textcomp}
\usepackage{microtype}
\usepackage[unicode=true]{hyperref}

\newcommand*{\mail}[1]{\href{mailto:#1}{\texttt{#1}}}
\newcommand*{\pkg}[1]{\textsf{#1}}

\addtokomafont{title}{\rmfamily}

\title{The \pkg{fnpara} package}
\subtitle{A \LaTeX\ style to typeset footnotes in run-on paragraphs.}
\author{Chris Rowley\and Dominik Wujastyk\and Bradley Smith}
\date{11/12/95}


\begin{document}

\maketitle

\section{History:}

A vain attempt at this \LaTeX\ modification was tried in late
September 1988 by Dominik Wujastyk, and sent to \TeX hax and UK\TeX\ %
for comment.  Chris Rowley came up with a working
version a few days later, but it still needs thorough testing.


\section{The Style:}

This \LaTeX\ style changes the way \LaTeX\ does footnotes.  With this
style invoked (\verb+\documentstyle[fnpara]{article}+), footnotes will
be typeset in a running paragraph, instead of one above another.
It is suitable for texts such as critical editions, which contain
many short footnotes.

This is based on the \TeX\ macros described by Knuth in the
\TeX book, Dirty Tricks, pages~398--400.  It is  not tested beyond
some simple examples.   In particular, it has not been tested
properly in connection with even the simplest of environments, nor with
other floats.

(In fact, ordinary LaTeX footnotes sometimes do not work well
if they need splitting (e.\,g.\ producing blank pages,
or appearing in shufled form): but these
problems should not affect this style.
I --CR-- have sent in a Bug Report about some
of these problems.)

Be aware of Knuth's note on the limitations of this method of
doing the job:  the \TeX\ stack is used four times per footnote,
and the stack is limited (\TeX book pp.~300--301).  So if you have
very many footnotes (in the hundreds) and encounter
%
\begin{verbatim}
"! TeX capacity exceeded, sorry [... save size ...]"
\end{verbatim}
%
errors, you may need to break your text into smaller sections.
Fortunately, this is very easy to do with \LaTeX.  And the
footnote counter can be reset to make the joins seamless.


\section{Authors}

\KOMAoption{parskip}{half}
\selectfont
Chris Rowley\\
Office tel: (England) 01 794 0575\\
Janet: \verb+CA_ROWLEY@UK.AC.OPEN.ACS.VAX+

Dominik Wujastyk\\
Office tel: (England) 01 387 4477 x 3013\\
Janet: \verb+Wujastyk@UK.AC.UCL.EUCLID+

October 3, 1988

\bigskip

Modified 11/12/95\\
Updated with new copies of relevant definitions from base files, and
added definitions for paragraph footnotes inside minipages

Bradley Smith\\
(612) 624--1535\\
\mail{basmith@prometheus.chem.umn.edu}

\end{document}
