\documentclass[a4paper]{article}
\usepackage{fixltx2e} % LaTeX patches, \textsubscript
\usepackage[T2A,T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[russian,english]{babel}
\usepackage[colorlinks=true,linkcolor=blue,urlcolor=blue]{hyperref}
\usepackage{bookmark}

\pagestyle{empty}

% Combine "TeX Gyre" fonts for Latin and "ParaType" fonts for Cyrillic:

% 1. Load paratype font-packages to get scaled Paratype fonts
\usepackage[scaled=0.9]{PTSerif}
\usepackage[scaled=0.9]{PTSans}
\usepackage[scaled=0.87]{PTMono}

% 2. Load TeX Gyre font-packages (after Paratype to use them as default fonts)
\usepackage{tgtermes}
\usepackage[scale=.90]{tgheros}
\usepackage{tgcursor}

% 3. Set up the substitutions:
\usepackage{substitutefont}
\substitutefont{T2A}{\rmdefault}{PTSerif-TLF}
\substitutefont{T2A}{\sfdefault}{PTSans-TLF}
\substitutefont{T2A}{\ttdefault}{PTMono-TLF}


\begin{document}

\section*{Cyrillic fonts to match Times/Helvetica/Curier}

The \href{http://www.ctan.org/pkg/tex-gyre}{TeX Gyre} project provides
extensions to the URW standard Postscript fonts. Cyrillic extensions by
Valek Filippov were part of TeXGyre but removed due to license
incompatibility. 

These fonts are still available in the 
\href{http://packages.debian.org/stable/scalable-cyrfonts-tex}%
{scalable-cyrfonts-tex} Debian package but not on CTAN.  Furthermore, both
\href{http://www.ctan.org/pkg/fouriernc}{Fourier New Century} and
\emph{Free New Century} use the TeX-name \emph{fnc}. Therefore
\emph{scalable-cyrfonts-tex} conflicts with the Debian
\href{http://packages.debian.org/stable/texlive-fonts-extra}%
{texlive-fonts-extra} package.

The Cyrillic fonts from the
\href{http://www.ctan.org/pkg/paratype}{paratype} package are not exact
matches but fit quite well if they are downscaled a bit. As the Paratype
package includes also T1 encoded fonts, Paratype fonts can be used for both
Latin and Cyrillic text.

Other vector fonts supporting Cyrillic include
\href{http://www.ctan.org/pkg/dejavu}{Deja Vu},
\href{http://www.ctan.org/pkg/droid}{Droid},
\href{http://www.ctan.org/pkg/gentium-tug}{Gentium}, and
\href{http://www.ctan.org/pkg/libertine}{Linux Libertine}.

\section*{\selectlanguage{russian}Тест шрифтов}

\newcommand{\teststring}{Hello world! 
	    \foreignlanguage{russian}{Здравствуй, мир!}}

\begin{description}
  \item [Normal:] \teststring

  \item [it:] \textit{\teststring}

  \item [sl:] \textsl{\teststring}%
      	      \footnote{TeX Gyre merges slanted and italic shapes}

  \item [bf:] \textbf{\teststring}

  \item [bf it:] \textbf{\textit\teststring}

  \item [sc:] not available in Paratype % \textsc{\teststring}

  \item [bf sc:] missing % \textsc{\textbf\teststring}

  \item [sf:] \textsf{\teststring}

  \item [sf bf:] \textsf{\textbf\teststring}

  \item [sf sl:] \textsf{\textsl\teststring}

  \item [sf sc:] not available in Paratype % \textsf{\textsc\teststring}

  \item [tt:] \texttt{\teststring}

\end{description}



\end{document}
