% An illustration of comment.sty by Victor Eijkhout,
% version 3.4 and later.
%
\documentclass{article}

\usepackage{comment}

\begin{document}

\includecomment{thisone}
\excludecomment{notthisone}

This sentence
\begin{comment}
has
\end{comment}
no verb.

Check for\begin{notthisone}
superfluous
\end{notthisone}
spaces (there should be none between `for' and `spaces'),
and around
\begin{thisone}
inclusion
\end{thisone}
of proper texts (one space between `around' and `inclusion',
and one between `inclusion' and `of').

Als check for
\begin{thisone}
multiple paragraphs.

Such as
\end{thisone}
here. (`Such as' should be a new paragraph.)

% a test of the special comments
\specialcomment{smallfry}{\begingroup\rmfamily\footnotesize}{\endgroup}
%\tracingmacros=2 \tracingcommands=2
This text is 
\begin{smallfry}
rather small
\end{smallfry}
don't you think? (The `rather small' should be footnotesize.)

And now we disable
\excludecomment{smallfry}
\begin{smallfry}
the special comment
\end{smallfry}
environment
(in between `disable' and `environment' is some excluded material).

% test of weird grouping
We switch to
\begin{thisone}
\bf bold face
\end{thisone}
inside \rm a comment. (The phrase `bold face inside' should be bold.)

\newcount\comlines
\specialcomment{countedcomment}
 {\comlines=0\relax \def\ProcessCutFile{}%
  \def\ThisComment##1{\global\advance\comlines1\relax}}
 {**Comment: \number\comlines\ line(s) removed**}
  
This is a line of text
\begin{countedcomment}
Oneline
\end{countedcomment}
another line of text
\begin{countedcomment}
One line
Two line
Three line
\end{countedcomment}
last line of text.

\specialcomment
    {underlinecomment}
    {%
      \def\ProcessCutFile{\input{\CommentCutFile}\relax}% Bug! this
                                % should not be necessary.
      \def\ThisComment##1{\WriteCommentLine{u: \underline{##1}\par}}
      \par
    }
    {\par}

A block of two underlined lines in between this line
\begin{underlinecomment}
line 1
line 2   
\end{underlinecomment}
and before this line.

\end{document}
