%% Ednotes.tex -- 
\title{\LARGE \bf \textit{ednotes.sty} \\ 
       for critical editions with \LaTeX \\[1ex]
       \normalfont \Large
       Lazy \texttt{ASCII}$\to$\texttt{PDF} listings}
\author{Uwe L\"uck}%%%\thanks{advised by Christian Tapp.}}
\documentclass[10pt]{article}
\usepackage{verbatim}
\pagestyle{headings}
\newcommand*{\cs}[1]{\texttt{\char"5C #1}}
\newcommand*{\lcurl}[1]{\[\texttt{CTAN:/macros/latex/contrib/#1}\]}
\begin{document}
\maketitle
\section*{Preface}

\texttt{ednotes[.sty]} is a macro package or bundle of 
macro files for scholarly critical editions---typed 
by U.\,L.\footnote{$\dots$ originally under the direction of 
Christian Tapp at a research project granted by the 
Deutsche Forschungsgemeinschaft (DFG).}
It is little more than a user interface for combining 
Stephan~I.\ B\"ottcher's \texttt{lineno.sty} and 
Alexander~I.~Rozhenko's \texttt{manyfoot}. 

Recent CTAN policy urges to upload \texttt{PDF} 
documentation accompanying \TeX\ macro files. 
% such files 
% do exist for the underlying \texttt{lineno.sty}--- 
% \lcurl{lineno.pdf} 
% (plus a supplementary \texttt{lnosuppl.pdf} there)---and 
% as 
% \lcurl{manyfoot.pdf}
% plus
% \lcurl{nccfoots.pdf}
% for the underlying \texttt{manyfoot.sty}\,. 
% 
% $\dots$ well, 
Sorry, a documentation of this kind doesn't exist 
at present for \texttt{ednotes} $\dots$---I am just offering 
some ersatz \texttt{PDF} documentation: listings of the 
\texttt{.sty} files belonging to the \texttt{ednotes} bundle, 
structured and navigatable through a table of contents. 

For making use of the options that \texttt{ednotes.sty} 
provides, it is vital to know the underlying 
\texttt{lineno.sty}---and its documentation. 
\texttt{ednotes} just passes \texttt{lineno} commands and 
options to the latter. So please notice 
\lcurl{lineno/lineno.pdf}
as well as (if you need it) 
\lcurl{lineno/lnosuppl.pdf\,.}
In the same directory, there are other macro files that 
\texttt{ednotes.sty} may need and whose documentation may 
be contained in \texttt{lnosuppl.pdf} (in the same 
pseudo-\texttt{PDF}-style as present). 

% Sorry, I haven't had the time for making a proper \texttt{PDF}
% documentation for the present \texttt{ednotes} bundle. 
% What I am offering is nothing but a somewhat structured listing 
% of the additional \texttt{.txt} and \texttt{.sty} files, 
% deriving from the \texttt{verbatim} package and its 
% \cs{verbatiminput} command. I hope that the high quality 
% (scalable) output is worth it. 

Please don't overlook that \lcurl{lineno/ednotugb.pdf} contains 
a (printable) \emph{overview} on the facilities of 
\texttt{ednotes}, as well as a comparison with Peter Wilson's 
alternative \texttt{ledmac}. 

% \clearpage 
\tableofcontents

\section{The \texttt{.txt} files}
\subsection{First encounter with \texttt{ednotes} on CTAN: 
            \texttt{README}}
\verbatiminput{README}
\subsection{Adding on installing etc.: \texttt{READMORE.txt}}
This is an extended "README" version 
for learning how to complete your download. 
\verbatiminput{READMORE.txt}
\subsection{What has changed since your earlier downloads: 
            \texttt{CHANGES.txt}}
\dots\ lists changes the bundle underwent in a chronolical order. 
\verbatiminput{CHANGES.txt}
\subsection{Advice for installation: \texttt{visible.txt}} 
Some very down-to-earth advices for real \TeX-newbies: 
\verbatiminput{visible.txt}
\subsection{Really recent? \texttt{CHANGING.txt}}
You might consider whether the snapshot (TeXLive most notably) 
of the \texttt{ednotes} bundle that you have received most 
recently displays the most recent (and perfect) state of it. 
\verbatiminput{CHANGING.txt}

\section{The main file: \texttt{ednotes.sty}}
\verbatiminput{ednotes.sty}

\section{Supplementary macro files}
\subsection{Tracking page break oscillations:
            \texttt{lblchng1.sty}}
\verbatiminput{lblchng1.sty}
\subsection{A lemma repeated in a line: \texttt{edcntwd0.sty}}
\verbatiminput{edcntwd0.sty}
\subsection{Some \texttt{para}-footnote hack: \texttt{mfparptc.sty}}
This deals with a problem that still needs much thinking \dots
\par
\verbatiminput{mfparptc.sty}
\subsection{A reminder of obsoleteness: \texttt{mfparxsp.sty}}
This file just tells that some earlier facility dealing with
vertical spacing has been superseded by \texttt{manyfoot}'s 
more recent \cs{ExtraParaSkip}\,.
\verbatiminput{mfparxsp.sty}

\section{\texttt{varnrule.tex}---customizing footnote rules}
\verbatiminput{varnrule.tex}

\end{document}

end of Ednotes.tex 
