\documentclass[draft]{exam-n}

\examdate{Today}
\examtime{Now! -- \emph{you're late!}}

\exambanner{Multiple-choice exam}

\degreedescriptions{Various degrees}
\schoolcoursecode{MCQ123}
\universitycoursecode{AssDes101}
\coursetitle{Assessment Design Strategies}

\rubric{Candidates should choose answers randomly where they feel
  unable to speculate wildly.}

\multiplechoiceanswers{4}

\begin{document}
\maketitle

\begin{mcq}
Which is the first answer in this list?
\answer Answer one
\item Answer two
\item Answer three
\item Answer four
\end{mcq}

\begin{mcq}
How many answers are in this list?
\item one
\item two
\item three
\answer four
\end{mcq}

\begin{mcq}
Which is the incorrect answer in this list?
\item Correct answer
\item Another correct answer
\answer The wrong answer
\item A final correct answer
\end{mcq}

\begin{mcq}
What is the nature of existence?
\item All good things come to those who wait
\item The universe makes sense
\answer Error is inevitable
\item Om
\end{mcq}

\begin{mcq}
How many roads must a man walk down?
\item Three
\answer Oooh, I wouldn't walk down there -- it's terribly windy.
\item Can I introduce you to the notion of the tuning fork?
\item Third left, then second right, and the nature of existence is
  available from the leprechaun in the little lane just after the bend
  in the road.
\end{mcq}

\begin{mcq}
How many questions are there in this test?
\item Six
\answer Eight
\item Ten
\item Twelve
\end{mcq}

\begin{mcq}
Is this a reasonable question?
\item no
\item yes
\item is this an answer?
\answer if you say so
\end{mcq}

\begin{mcq}
What is the most important thing about examinations?
\item They are an accurate measure of understanding
\item They are repeatable and reliable
\answer They are nicely formatted
\item They are stress-free occasions which everyone enjoys.
\end{mcq}

\end{document}
