\documentclass[oldtoc,a4paper,10pt]{artikel3}
\usepackage{a4wide}
\begin{document}
\author{J.B.~Rhebergen}
\date{\today}
\title{{\tt sig.sty} Documentation}
\maketitle

This is the {\tt sig.sty} package which can be used to put a scanned
signature in your fax document or other documents. If you want to use this
{\tt sig.sty} there are a number of things you need to change.

The dimension of the \verb|\vbox| and \verb|\hbox| depend on the size of
the signature. Measure them and change accordingly. The \verb|\sig| command
puts the signature in the text. I added some negative kern which causes the
signature to overlap a little just as happens sometimes when you put your
signature on a document by hand. If you don't want this, use
\verb|\altsig|. The \verb|\zsig| and \verb|zaltsig| are similar except that
the \verb|\vbox| and \verb|\hbox| have zero dimensions. This allows you to
put the signature on top of other text. (yes I know that is weird but it
can be handy sometimes)

Remember that you need to scale your signature to the proper size with
a graphics program. To use with a fax document in fax resolution just put
the command \verb|\usepackage[fax]{sig}| in the preamble of your document.
In non fax document leave out the {\tt fax} option.

\footnotesize
\begin{verbatim}
\NeedsTeXFormat{LaTeX2e}
\ProvidesPackage{sig}[1995/07/04 v1.0 by: JBR]

\newcommand{\sig}{\leavevmode
  \hbox to 62mm{\kern-5mm\vbox to 12mm{\kern-5mm\special{em:graph sig}\vfil}\hfil}}
\newcommand{\altsig}{\leavevmode
  \hbox to 67mm{\vbox to 17mm{\special{em:graph sig}\vfil}\hfil}}
\newcommand{\zsig}{\leavevmode
  \hbox to 0mm{\kern-5mm\vbox to 0mm{\kern-10mm\special{em:graph sig}\vfil}\hfil}}
\newcommand{\zaltsig}{\leavevmode
  \hbox to 0mm{\vbox to 0mm{\special{em:graph sig}\vfil}\hfil}}
                   
\DeclareOption{fax}{\typeout{Using signature for fax!}
\renewcommand{\sig}{\leavevmode%
  \hbox to 62mm{\kern-5mm\vbox to 12mm{\kern-5mm\special{em:graph faxsig}\vfil}\hfil}}
\renewcommand{\altsig}{\leavevmode%
  \hbox to 67mm{\vbox to 17mm{\special{em:graph faxsig}\vfil}\hfil}}
\renewcommand{\zsig}{\leavevmode%
  \hbox to 0mm{\kern-5mm\vbox to 0mm{\kern-10mm\special{em:graph faxsig}\vfil}\hfil}}
\renewcommand{\zaltsig}{\leavevmode%
  \hbox to 0mm{\vbox to 0mm{\special{em:graph faxsig}\vfil}\hfil}}
}
\ProcessOptions
\end{verbatim}
\normalsize

\end{document}
