\documentclass[10pt,letterpaper,busletter,dateno,addrfromright,addrfromleft,orderfromtodate,%
faxheaderpage]{newlfm}

\lthmastg\setadrto{\adrPIT}\sigsLvBa

\leftmarginsize{.75in}
\rightmarginsize{1.385in}
\dateskipbefore{-10pt}
\dateskipafter{.45in}
\addrfromskipafter{35pt}
\bottommarginskip{5pt}
\bottommarginskipbelow{-5pt}
\addrtoskipafter{13pt}
\greettoskipafter{.25in}
\addrfromskipbefore{0pt}
\topmarginskip{3pt}

\rheader{New header text!!}
\lheader{The Font Institute --- Propaganda Release 3993921.2}

\closeline{Yours in fontal health,}
\Cfooter{\parbox{3in}{{\Large\bfseries\itshape Your font is my font!!}}}

\PhrRegard{About}\regarding{Fonts and typography}
\greetto{Tovarich Pyotr:}

\begin{document}
\begin{newlfm}
\showdim{In body of text}
In terms of the relationships between font descriptions and musical
typography, it seems to me that only in the case of the early 1800s can even
the slightest case be made for a relationship.  

\hyphenation{ble-eeee-ah}

1) Natter, blather, squiggle, gnorp, bleeeeeah.
2) Natter, blather, squiggle, gnorp, bleeeeeah.
3) Natter, blather, squiggle, gnorp, bleeeeeah.
4) Natter, blather, squiggle, gnorp, bleeeeeah.
5) Natter, blather, squiggle, gnorp, bleeeeeah.
6) Natter, blather, squiggle, gnorp, bleeeeeah.
7) Natter, blather, squiggle, gnorp, bleeeeeah.
8) Natter, blather, squiggle, gnorp, bleeeeeah.
9) Natter, blather, squiggle, gnorp, bleeeeeah.
10) Natter, blather, squiggle, gnorp, bleeeeeah.
11) Natter, blather, squiggle, gnorp, bleeeeeah.
12) Natter, blather, squiggle, gnorp, bleeeeeah.
13) Natter, blather, squiggle, gnorp, bleeeeeah.
14) Natter, blather, squiggle, gnorp, bleeeeeah.
15) Natter, blather, squiggle, gnorp, bleeeeeah.
16) Natter, blather, squiggle, gnorp, bleeeeeah.
17) Natter, blather, squiggle, gnorp, bleeeeeah.
18) Natter, blather, squiggle, gnorp, bleeeeeah.
19) Natter, blather, squiggle, gnorp, bleeeeeah.
20) Natter, blather, squiggle, gnorp, bleeeeeah.
21) Natter, blather, squiggle, gnorp, bleeeeeah.
22) Natter, blather, squiggle, gnorp, bleeeeeah.
23) Natter, blather, squiggle, gnorp, bleeeeeah.
24) Natter, blather, squiggle, gnorp, bleeeeeah.
25) Natter, blather, squiggle, gnorp, bleeeeeah.
26) Natter, blather, squiggle, gnorp, bleeeeeah.
27) Natter, blather, squiggle, gnorp, bleeeeeah.
28) Natter, blather, squiggle, gnorp, bleeeeeah.
29) Natter, blather, squiggle, gnorp, bleeeeeah.
30) Natter, blather, squiggle, gnorp, bleeeeeah.
31) Natter, blather, squiggle, gnorp, bleeeeeah.
32) Natter, blather, squiggle, gnorp, bleeeeeah.
33) Natter, blather, squiggle, gnorp, bleeeeeah.
34) Natter, blather, squiggle, gnorp, bleeeeeah.
35) Natter, blather, squiggle, gnorp, bleeeeeah.
36) Natter, blather, squiggle, gnorp, bleeeeeah.
37) Natter, blather, squiggle, gnorp, bleeeeeah.
38) Natter, blather, squiggle, gnorp, bleeeeeah.
39) Natter, blather, squiggle, gnorp, bleeeeeah.
40) Natter, blather, squiggle, gnorp, bleeeeeah.
41) Natter, blather, squiggle, gnorp, bleeeeeah.
42) Natter, blather, squiggle, gnorp, bleeeeeah.
43) Natter, blather, squiggle, gnorp, bleeeeeah.
44) Natter, blather, squiggle, gnorp, bleeeeeah.
45) Natter, blather, squiggle, gnorp, bleeeeeah.
46) Natter, blather, squiggle, gnorp, bleeeeeah.
47) Natter, blather, squiggle, gnorp, bleeeeeah.
48) Natter, blather, squiggle, gnorp, bleeeeeah.
49) Natter, blather, squiggle, gnorp, bleeeeeah.
50) Natter, blather, squiggle, gnorp, bleeeeeah.
51) Natter, blather, squiggle, gnorp, bleeeeeah.
52) Natter, blather, squiggle, gnorp, bleeeeeah.
53) Natter, blather, squiggle, gnorp, bleeeeeah.
54) Natter, blather, squiggle, gnorp, bleeeeeah.
55) Natter, blather, squiggle, gnorp, bleeeeeah.
56) Natter, blather, squiggle, gnorp, bleeeeeah.
57) Natter, blather, squiggle, gnorp, bleeeeeah.
58) Natter, blather, squiggle, gnorp, bleeeeeah.
59) Natter, blather, squiggle, gnorp, bleeeeeah.
60) Natter, blather, squiggle, gnorp, bleeeeeah.
61) Natter, blather, squiggle, gnorp, bleeeeeah.
62) Natter, blather, squiggle, gnorp, bleeeeeah.
63) Natter, blather, squiggle, gnorp, bleeeeeah.
64) Natter, blather, squiggle, gnorp, bleeeeeah.
65) Natter, blather, squiggle, gnorp, bleeeeeah.
66) Natter, blather, squiggle, gnorp, bleeeeeah.
67) Natter, blather, squiggle, gnorp, bleeeeeah.
68) Natter, blather, squiggle, gnorp, bleeeeeah.
69) Natter, blather, squiggle, gnorp, bleeeeeah.
70) Natter, blather, squiggle, gnorp, bleeeeeah.
71) Natter, blather, squiggle, gnorp, bleeeeeah.
72) Natter, blather, squiggle, gnorp, bleeeeeah.
73) Natter, blather, squiggle, gnorp, bleeeeeah.
74) Natter, blather, squiggle, gnorp, bleeeeeah.
75) Natter, blather, squiggle, gnorp, bleeeeeah.
76) Natter, blather, squiggle, gnorp, bleeeeeah.
77) Natter, blather, squiggle, gnorp, bleeeeeah.
78) Natter, blather, squiggle, gnorp, bleeeeeah.
79) Natter, blather, squiggle, gnorp, bleeeeeah.
80) Natter, blather, squiggle, gnorp, bleeeeeah.
81) Natter, blather, squiggle, gnorp, bleeeeeah.
82) Natter, blather, squiggle, gnorp, bleeeeeah.
83) Natter, blather, squiggle, gnorp, bleeeeeah.
84) Natter, blather, squiggle, gnorp, bleeeeeah.
85) Natter, blather, squiggle, gnorp, bleeeeeah.
86) Natter, blather, squiggle, gnorp, bleeeeeah.
87) Natter, blather, squiggle, gnorp, bleeeeeah.
88) Natter, blather, squiggle, gnorp, bleeeeeah.
89) Natter, blather, squiggle, gnorp, bleeeeeah.
90) Natter, blather, squiggle, gnorp, bleeeeeah.
91) Natter, blather, squiggle, gnorp, bleeeeeah.
92) Natter, blather, squiggle, gnorp, bleeeeeah.
93) Natter, blather, squiggle, gnorp, bleeeeeah.
94) Natter, blather, squiggle, gnorp, bleeeeeah.
95) Natter, blather, squiggle, gnorp, bleeeeeah.
96) Natter, blather, squiggle, gnorp, bleeeeeah.
97) Natter, blather, squiggle, gnorp, bleeeeeah.
98) Natter, blather, squiggle, gnorp, bleeeeeah.
99) Natter, blather, squiggle, gnorp, bleeeeeah.
100) Natter, blather, squiggle, gnorp, bleeeeeah.
101) Natter, blather, squiggle, gnorp, bleeeeeah.
102) Natter, blather, squiggle, gnorp, bleeeeeah.
103) Natter, blather, squiggle, gnorp, bleeeeeah.
104) Natter, blather, squiggle, gnorp, bleeeeeah.
105) Natter, blather, squiggle, gnorp, bleeeeeah.
106) Natter, blather, squiggle, gnorp, bleeeeeah.
107) Natter, blather, squiggle, gnorp, bleeeeeah.
108) Natter, blather, squiggle, gnorp, bleeeeeah.
109) Natter, blather, squiggle, gnorp, bleeeeeah.
110) Natter, blather, squiggle, gnorp, bleeeeeah.
111) Natter, blather, squiggle, gnorp, bleeeeeah.
112) Natter, blather, squiggle, gnorp, bleeeeeah.
113) Natter, blather, squiggle, gnorp, bleeeeeah.
114) Natter, blather, squiggle, gnorp, bleeeeeah.
115) Natter, blather, squiggle, gnorp, bleeeeeah.
116) Natter, blather, squiggle, gnorp, bleeeeeah.
117) Natter, blather, squiggle, gnorp, bleeeeeah.
118) Natter, blather, squiggle, gnorp, bleeeeeah.
119) Natter, blather, squiggle, gnorp, bleeeeeah.
120) Natter, blather, squiggle, gnorp, bleeeeeah.
121) Natter, blather, squiggle, gnorp, bleeeeeah.
122) Natter, blather, squiggle, gnorp, bleeeeeah.
123) Natter, blather, squiggle, gnorp, bleeeeeah.
124) Natter, blather, squiggle, gnorp, bleeeeeah.
125) Natter, blather, squiggle, gnorp, bleeeeeah.
126) Natter, blather, squiggle, gnorp, bleeeeeah.
127) Natter, blather, squiggle, gnorp, bleeeeeah.
128) Natter, blather, squiggle, gnorp, bleeeeeah.
129) Natter, blather, squiggle, gnorp, bleeeeeah.
130) Natter, blather, squiggle, gnorp, bleeeeeah.
131) Natter, blather, squiggle, gnorp, bleeeeeah.
132) Natter, blather, squiggle, gnorp, bleeeeeah.
133) Natter, blather, squiggle, gnorp, bleeeeeah.
134) Natter, blather, squiggle, gnorp, bleeeeeah.
135) Natter, blather, squiggle, gnorp, bleeeeeah.
136) Natter, blather, squiggle, gnorp, bleeeeeah.
137) Natter, blather, squiggle, gnorp, bleeeeeah.
138) Natter, blather, squiggle, gnorp, bleeeeeah.
139) Natter, blather, squiggle, gnorp, bleeeeeah.
140) Natter, blather, squiggle, gnorp, bleeeeeah.
141) Natter, blather, squiggle, gnorp, bleeeeeah.
142) Natter, blather, squiggle, gnorp, bleeeeeah.
143) Natter, blather, squiggle, gnorp, bleeeeeah.
144) Natter, blather, squiggle, gnorp, bleeeeeah.
145) Natter, blather, squiggle, gnorp, bleeeeeah.
146) Natter, blather, squiggle, gnorp, bleeeeeah.
147) Natter, blather, squiggle, gnorp, bleeeeeah.
148) Natter, blather, squiggle, gnorp, bleeeeeah.
149) Natter, blather, squiggle, gnorp, bleeeeeah.
150) Natter, blather, squiggle, gnorp, bleeeeeah.

In conclusion, I have little to say on the matter.
\end{newlfm}
\end{document}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: t
%%% End: 
