\RequirePackage{flashmovie}
% it is neccessay to use "\RequirePackage{flashmovie}" because beamer
% also uses "\pdfminorversion". see flashmovie.sty for an explanation.

%---------------------------------------------
% pure-flash-example
%---------------------------------------------

\documentclass[10pt]{beamer}

\usepackage[english]{babel}
\usepackage{hyperref}

\usetheme{Warsaw} % Warsaw,Hannover,boxes
\usecolortheme{rose} % orchid,lily,dolphin,beetle,crane
\usefonttheme{professionalfonts} % professionalfonts,serif
\useinnertheme{rounded} % rounded
\useoutertheme{shadow} % shadow,sidebar,split

%---------------------------------------------

\setbeamersize{text margin left=0.3cm}   % small margins
\setbeamersize{text margin right=0.3cm}

%---------------------------------------------
% for writing a document is is convenient to switch movies off. to do this,
% use "blank=1" as an option. in order to use the videos, use "blank=0".
% 
% to make thinks easy, use "blank=\myblank" as an option and define
% \myblank as suitable.

\def\myblank{0}
%\def\myblank{1}

%---------------------------------------------

\title[flashmovie.sty]{flashmovie.sty}
\author[\href{mailto:thartmann15@googlemail.com}{Timo Hartmann}]{\href{mailto:thartmann15@googlemail.com}{Timo Hartmann}}

\date[2010]{}

%---------------------------------------------

\begin{document}

\frame{

\titlepage

\begin{block}{\alert{Warning}}
It is recommended to use the latest available version of the Adobe Reader
to view PDF files generated with flashmovie.sty.
% Otherwise your Adobe Reader may die a sudden painfull death...
\end{block}

} % end frame

%---------------------------------------------

\begin{frame}

\frametitle{Basics}

\begin{minipage}[t]{6.1cm}
\vspace{0cm}
\flashmovie[width=6cm,height=5cm]{saturn5.swf}
\end{minipage}
\begin{minipage}[t]{5.7cm}
\vspace{0cm}
This package allows direct embedding of flash movies into PDF files. It is
designed for use with pdflatex.
\vspace{0.5cm}

Basically it uses the fact that the Adobe Reader 9 contains an embedded Adobe Flash 
player which can be invoked with the ``rich media annotation'' feature which is described 
in ``Adobe Supplement to the ISO 32000 BaseVersion: 1.7 ExtensionLevel: 3''.
\vspace{0.5cm}

This means that you can only use flashmovie.sty in combination
with Adobe Reader 9 and upwards.

\end{minipage}

\vspace{0.5cm}

P.S.: This sample video is courtesy of the NASA ( \href{http://heasarc.gsfc.nasa.gov/Videos/historical/saturn5.avi}{saturn5.avi} ).

\end{frame}

%---------------------------------------------

\begin{frame}[fragile]

\frametitle{Explanation}

The source code used for the video on the previous page is:

\begin{verbatim}
\flashmovie[width=6cm,height=5cm]{saturn5.swf}
\end{verbatim}

This means that the movie is really a flash animation and is not
rendered by a player. So no controls are available.

\end{frame}

%---------------------------------------------

\begin{frame}[fragile]

\frametitle{Issues with the flv-player}

\begin{itemize}

\item Sadly the flv-player does not work reliably with the latex package
beamer. If you try it, the acrobat reader often crashes while trying
to change the page. I have no idea what is the cause of this problem.

\item The only reliable way to use beamer seems to be directly embedding the videos as
      flash animations or to use the JW player.

\end{itemize}

\end{frame}

%---------------------------------------------

\begin{frame}[fragile]

\frametitle{Advanced use}

Rich media annotations are not restricted to videos.
You can use flash animations, too.
Here for example is a clock written in action script:

\flashmovie[width=8cm,height=5cm,blank=\myblank]{clock.swf}

\begin{verbatim}
  \flashmovie[width=8cm,height=5cm]{clock.swf}
\end{verbatim}

\end{frame}

%---------------------------------------------

\end{document}
