\documentclass{article}
\usepackage{amsmath}
%
% the driver line is not necessary if you
% have aebpro.cfg configured to your driver.
%
\usepackage[%
    web={designv,pro,tight,usesf,usetemplates},
    eforms,graphicxsp={showembeds},aebxmp
]{aeb_pro}
\usepackage{yt4pdf}
\usepackage{picins}

% \textBgColor{lightgray}

\DeclareDocInfo
{
    title=The \texorpdfstring{\textsf{yt4pdf} Package\\[3pt]}{yt4pdf Package: }Demo 2 of You Tube API,
    author=D. P. Story,
    university=Acro\negthinspace\TeX.Net,
    email=dpstory@acrotex.net,
    subject={Live streamed video from YouTube.com},
    keywords={You Tube, AeB, rmannot},
    talksite=\url{http://www.acrotex.net},
    talkdate={\today}
}
\talkdateLabel{Published:}

\def\AcroTeX{Acro\!\TeX}

\makePoster{pdfyt_poster}{pdfyt_poster}
\makePoster{yt_poster}{graphics/yt_icon_sm}

\parindent=0pt\parskip6pt
\rheader{}

\DeclarePageLayout{footskip=15.5pt}

%\previewtrue

\declarePlayList{\playList}{%
    \ytIdTitle{Kung-Fu Fighting {(Bruce Lee version)}}{GZ9e3Dy7obA}
    \ytIdTitle{Kung-Fu Fighting {(Original, must view on YouTube)}}{jhUkGIsKvn0}
    \ytIdTitle{RAF Sculthorpe {(Watch for my Mom!)}}{K7H8QghVs9k}
    \ytIdTitle{Rocket Jump}{7XzdZ4KcI8Y}
}

%
% Note: popupmenu package version 1.0 (2010/09/18) or later required here.
%
\begin{popupmenu}{YTMenu}
    \puIdTitle{Select a You Tube Video}{} % A title has no yt Id
    \begin{submenu}{title=Music Videos}
        \puIdTitle{Kung-Fu Fighting (Bruce Lee version)}{GZ9e3Dy7obA}
        \puIdTitle{Kung-Fu Fighting (Original, view on You Tube)}{jhUkGIsKvn0}
        \puIdTitle{Ink Spots - If I didn't care}{rvwfLe6sLis}
        \puIdTitle{Ink Spots - My Prayer}{h7KJCns5v3g}
    \end{submenu}
    \begin{submenu}{title=Adobe Related Videos}
        \puIdTitle{Open at Adobe}{eNzrn8-JFSE}
        \puIdTitle{Lori's Corner Episode \#1}{5y9-EVmreU4}
        \puIdTitle{Intro to PDF Portfolios}{T9Yzo_hlwP0}
    \end{submenu}
    \begin{submenu}{title=Action Videos}
        \puIdTitle{Rocket Jump}{7XzdZ4KcI8Y}
        \puIdTitle{AimBot}{NYGlWjIKoY4}
        \puIdTitle{Flower Warfare}{031Dshcnso4}
        \puIdTitle{Modern Warfare: Frozen Crossing Pt. 1}{rzQMGMuD1sU}
        \puIdTitle{Modern Warfare: Frozen Crossing Pt. 2}{a08ViuvORRM}
    \end{submenu}
\end{popupmenu}

% After defining all menus used with YouTube videos, we declare them using \ytPopupMenuData
% The argument is a token list: \ytPopupMenuData{\myEducationMenu\myFunMenu...}
\ytPopupMenuData{\YTMenu}

\begin{document}

\maketitle

\input{config_intacc.def}


\begin{center}\bfseries\Large\color{blue}YT4PDF\end{center}

\DeclareYTName{ytInWindow}
\ytBtnPresets{\BC{.5 0 0}\W{1}\S{S}\textFont{HeBo}}%
{\setlength{\dimen0}{(\linewidth)/6-1bp}\xdef\ytWidth{\the\dimen0}} %-.5bp
\cfooter{\makebox[\linewidth][c]{\kern1bp\ytPlayToggle{\ytWidth}{11bp}\kern1bp\ytStopLoading[\textSize{0}]{\ytWidth}{11bp}\kern1bp
\ytMuteToggle{\ytWidth}{11bp}\kern1bp\ytRewind{\ytWidth}{11bp}\kern1bp
\ytSelect[\textSize{0}]{eNzrn8-JFSE}{\ytWidth}{11bp}\kern1bp\ytOpenWeb[\textSize{0}]{\ytWidth}{11bp}}}

\newcommand{\myPBPresets}{\CA{YT Menu}\textColor{0 0 1}\W1\BC{}\textSize{0}
    \S{S}\AA{\AAMouseEnter{\JS{ytPopupMenu("\ytCurrRMAName",YTMenu)}}}}

\setWindowDimPos{position={halign=center,valign=center,hoffset=0,voffset=0},
    width={default=571},height={default=330}}%

\setlength{\pichskip}{3pt}
\parpic(.25in,.25in){%
\parbox{.25in}{\offinterlineskip\resizebox{.25in}{!}
{\ytRmAnnot[windowed,poster=yt_poster,deactivated=pageclose]{ytInWindow}{72bp}{72bp}}}}
\pushButton[\presets{\myPBPresets}]{myYTMenu}{20bp}{5bp}
On this page, the You Tube annotation is icon size and the video is played
in a floating window.

Move the window by dragging its title bar. Resize it by dragging the
lower-right corner. Close the window by clicking on the upper-right
corner.

The \textsf{YT Menu} uses the \textsf{popupmenu} package. Alternatively,
we could just use the combo box that has been used already, but, I thought
we might try something different.

\newpage

\ytRmAnnot[posternote=AcroTeX PDF Tube,deactivated=pageclose]{YouTube1}{571bp/2}{330bp/2}\\
\ytPlayList{GZ9e3Dy7obA}{\playList}\ytComboListPresets{\textSize{9}}%
\ytComboList{(571bp/2)-1bp-33bp}{11bp}\kern1bp\ytComboBtn[\TU{Click to play, shift-click to load thumbnail only}]{33bp}{11bp}

\newpage
\cfooter{}
\begin{minipage}{571bp/2}
\ytRmAnnot[posternote=AcroTeX PDF Tube,deactivated=pageclose]{YouTube2}{571bp/2}{330bp/2}\\
\ytPlayList{GZ9e3Dy7obA}{\playList}\ytComboListPresets{\textSize{9}}%
\ytComboList{(571bp/2)-1bp-33bp}{11bp}\kern1bp\ytComboBtn[\TU{Click to play, shift-click to load thumbnail only}]{33bp}{11bp}
\end{minipage}\hfill
\begin{minipage}{\linewidth-290bp-10bp}\centering
\ytPlayToggle{1in}{11bp}\\[3bp]
\ytStopLoading[\textSize{0}]{1in}{11bp}\\[3bp]
\ytMuteToggle{1in}{11bp}\\[3bp]
\ytRewind{1in}{11bp}\\[3bp]
\ytSelect[\textSize{0}]{eNzrn8-JFSE}{1in}{11bp}\\[6bp]
\ytOpenWeb[\textSize{0}]{1in}{11bp}
\end{minipage}

\newpage

\begin{center}
\begin{minipage}{571bp/2}
\ytRmAnnot[poster=pdfyt_poster,deactivated=pageclose]{YouTube3}{571bp/2}{330bp/2}\\
\ytPlayList{GZ9e3Dy7obA}{\playList}\ytComboListPresets{\textSize{9}}%
\ytComboList{(571bp/2)-1bp-33bp}{11bp}\kern1bp\ytComboBtn[\TU{Click to play, shift-click to load thumbnail only}]{33bp}{11bp}\\[1bp]
{\setlength{\dimen0}{(\linewidth+1bp)/6-1bp}\xdef\ytWidth{\the\dimen0}}%
\ytPlayToggle{\ytWidth}{11bp}\kern1bp\ytStopLoading[\textSize{0}]{\ytWidth}{11bp}\kern1bp
\ytMuteToggle{\ytWidth}{11bp}\kern1bp\ytRewind{\ytWidth}{11bp}\kern1bp
\ytSelect[\textSize{0}]{eNzrn8-JFSE}{\ytWidth}{11bp}\kern1bp
\ytOpenWeb[\textSize{0}\CA{View on Web}]{\ytWidth}{11bp}
\end{minipage}
\end{center}

\end{document}
