\documentclass[a4paper]{article}

\usepackage{tabularx}

%\usepackage{doublespace}
%\setstretch{1.2}

\usepackage{ae}
\usepackage[T1]{fontenc}
\usepackage{CV}

\begin{document}

\pagestyle{empty}

%Ueberschrift
\begin{center}
\huge{\textsc{Curriculum Vitae}}
\vspace{\baselineskip}

\Large{\textsc{Philipp Maier}}
\end{center}
\vspace{1.5\baselineskip}

\section{Address}

\begin{flushleft}
  Steenhouwerskade 12345 \\
  9718 DL G-Town \\
  The Netherlands \\
  Phone: +31-50-313.7020 \\
  Fax: +31-50-363.7337 \\
  Email: pmaier@yahoo.com \\
  Homepage: www.philipp-maier.de\\
\end{flushleft}


\section{Personal Details}
\begin{flushleft}
  Gender: Male \\
  Date of birth: 36th of November, 1971 \\
  Place of birth: R-Town, Germany \\
  Present Citizenship: German
\end{flushleft}


\section{Education}

\begin{CV}
\item[10/1992--09/1994] Undergraduate Studies in General Economics at
  the University of S-City, Germany
\item[10/1994--04/1997] M.A. in General Economics at the University of
  G-Town, Germany

Specialization: Macroeconomics, Monetary Economics, Public Finance and Econometrics

Thesis: \emph{Adverse selection and signaling}; Supervisor: Prof. Dr. V. Smart.
\item[Since 07/1997] Ph.D. student at the University of G-Town,
  The Netherlands.

  Project title: \emph{The position of the Bundesbank in international perspective};
  Supervisor: Prof. Dr. P.Fessor

  I intend to complete the Ph.D. in March, 2001.
\end{CV}

\pagebreak

\section{Thesis}
\noindent A special feature of the German financial system is the position of
the \emph{Deutsche Bundesbank}, which is considered to be one of the
most independent central banks in the world. The purpose of this
research is to analyse to which extent this formal independent
position has led to less political influence on the central
bank than, for instance, in the USA.

\section{Working Experience}

\begin{CV}

\item[09/1991--10/1992] Civil Service: Working in a
  kindergarden in C-Town, Germany

  Alternative for a military service: This was a special kindergarden for children from
  socially challenged families.

\item[02/1991--10/1992] Reporting work for the daily newspaper
  `\emph{ABC News}' in Z-Town, Germany (part-time)

\item[03/1995--04/1995] Internship at the Office for Regional
  Statistics in Y-Town, Germany

  \emph{Overview:} The construction of new motorways is often
  considered as a means to foster economic growth in less developed
  regions. This research project aimed at investigating the economic
  effects of newly build motorways in Germany. My main tasks were to
  provide an overview over the literature and to develop the basic
  structure of a model which was later used to to quantify the
  economic effects.


\item[03/1995--03/1997] Research Assistant at the Faculty of
  Economics, University of G-Town, Germany (part-time)

\end{CV}


\section{Language Knowledge}
\begin{table}[h] %\centering
\begin{tabular}{p{2cm}>{\bfseries}p{2.5cm}p{3cm}}
& German  & native \\
& French  & near native \\
& Dutch & near native\\
& Italian & fair \\
\end{tabular}
\end{table}

\pagebreak

\section{References}

\noindent These persons are familiar with my professional qualifications and my character:

\begin{table}[h]
\begin{tabular}{@{}lll@{}}
\textbf{Prof. Dr. Pro. Fessor} \\
Thesis supervisor & Phone: & +31-50-312.3456\\
P.O. Box 800 & Fax: & +31-50-567.123\\
9700 AV Groningen & Email: & p.fessor@xxx.rug.nl \\
The Netherlands \\
\end{tabular}
\end{table}

\vspace{2\baselineskip}
\noindent Groningen, \today



\end{document}

%Tabellen
\begin{table}[htbp] \centering%
\begin{tabular}{lll}\hline\hline
1 & 2 & 3 \\ \hline
1 & \multicolumn{2}{c}{2} \\
\hline
\end{tabular}
\caption{Titel\label{Tabelle: Label}}
\end{table}






