\documentclass{article}
\usepackage[T1]{fontenc}
\usepackage[ansinew]{inputenc}
\usepackage[english]{babel}
\usepackage{csquotes}
\usepackage[section,toc1,hyperref]{javadoc}

\hypersetup{colorlinks,citecolor=black,filecolor=black,linkcolor=black,urlcolor=black}

\author{Jolle\footnote{Comments, Help, Questions, Critics to \mbox{joerman.lieder@gmx.net}}  }
\date{\today}
\title{Documentation of TexGen-Doclet}

\begin{document}
\maketitle

\abstract{This documentation describes the use of the TexGen-Doclet. A doclet is a class derived from com.sun.javadoc.Doclet that can be used to generate a documentation with the javadoc-tool out of commented java-source-code. By default javadoc generates a HTML documentation, TexGen generates \TeX-files according to the javadoc-package. The doclet is under GNU GENERAL PUBLIC LICENSE\footnote{www.gnu.org}}

\tableofcontents

\section{Usage}
\subsection{Calling javadoc}
To use the doclet it has to be a parameter for the javadoc programm. The javadoc syntax is:

javadoc [options] [packagenames] [sourcefiles] [@files]

Possible Options are -doclet and -docletpath. For TexGen write \enquote{TeXGen} after -doclet and after -docletpath the Path to the TexGen.jar jararchive.

\subsection{Option for TexGen}
TexGen provides an own option. You can use -dest and the path to the outputfolder, where the generated files will be located. If you don't use this option, an folder named \enquote{texgendoc} will be created.

The final syntax is:

javadoc -doclet TexGen -docletpath <...texgen.jar> -dest <ausgabepath>  ...

\subsection{Using Eclipse}
With Eclipse Javadoc can be called with an plugin. You have the possibility to choose a custom doclet. Use TexGen for the docletname and the path to the jararchive for the docletpath. The next dialog provides the inputfield for extra javadoc options. Here you can add -dest <outputpath>.

\section{Known and Open Issues}
\begin{itemize}
	\item Special character in the documented text are converted not to interprete them as Tex-control character. These are:
\verb+{ } _ ^ & # [ ]+. All other special characters, that might build a tex-control-character should be avoided.
	\item Generic data types aren't supported.
	\item Duplicate Classnames (e.g. in different packages) aren't supported in linking.
\end{itemize}

\section{Source Code documentation}

\input{doc/TexGen}
\input{doc/ClassWriter}
\input{doc/TexPrintStream}
\input{doc/InhTable}
\input{doc/InhEntry}


\end{document}