\documentclass[11pt]{book}
\usepackage[repeat]{poetry}
\usepackage[pagestyles,innermarks,clearempty]{titlesec}[2005/01/22 v2.6]
\usepackage{example}

\TextHeight {6.8in}
\TextWidth  {4.25in}

\hyphenation {chi-ca-go mo-not-o-nous}
\hfuzz 1pt

\Novus \titulus \GroupTitle
\Facies         {\RelSize{2}\textsl{#1}}
\SpatiumSupra   {\penalty -10000} 
\SpatiumInfra   {2\leading plus 1\leading \penalty 10000}

\newenvironment {poem} [1]
 {\titulus{#1}%
  \versus
  \Nstanza{0}%
  \numerus{1}}
 {\endversus}

\Novus \numerus \Nstanza
\Facies         {\RelSize{+1}\textup{#1}}

\Novus \titulus \Snumber
\Facies         {\Nstanza*{=+1}}
\SpatiumSupra   {1\leading plus 1\leading minus .5\leading \penalty -50}
\SpatiumInfra   {.25\leading plus .125\leading \penalty 10000}

\Facies \tituli {\RelSize{+1}\textbf{#1}}
\SpatiumSupra   {2\leading plus 1\leading minus .5\leading \penalty -50}
\SpatiumInfra   {.5\leading plus .25\leading \penalty 10000}

\Locus \excessus {\textleftmargin + 4em}
\Modus           {\unhyphenated}

\Locus \numeri   {\rightmargin +3em}
\Facies          {\RelSize{-1}\oldstylenums{#1}}


\newpagestyle {whitman} {
  \sethead   [\oldstylenums{\thepage}]
             [\textsc{walt whitman}]
             []
             {}
             {\textsc{memories of president lincoln}}
             {\oldstylenums{\thepage}}
  \Capita    {section}{subsection}
}

\Novum \spatium \vacua
\Facies         {.5\leading plus .25\leading minus .125\leading}


\Novum \spatium \Vacua
\Facies         {1\leading plus .5\leading minus .25\leading}

\begin{document}

\ExampleTitle {Walt Whitman}
              {Leaves of Grass\\[3ex]
              \RelSize {-2}\textit{Memories of President Lincoln}}
              {Signet Classics\\[.5ex]New American Library, 1958}

\pagestyle {whitman} 
\thispagestyle {empty} 


\GroupTitle {MEMORIES OF PRESIDENT LINCOLN}


\begin{poem}{When Lilacs Last in the Dooryard Bloom'd}


  \Modus \excessus {\justified[850]}
  \Facies \strophae {\Snumber}
  \Forma            {0{\penalty 1000}0}

\stropha % 1

When lilacs last in the dooryard bloom'd,
And the great star early droop'd in the western sky in the night,
I mourn'd, and yet shall mourn with ever-returning spring.
     \vacua
Ever-returning spring, trinity sure to me you bring,
Lilac blooming perennial and drooping star in the west,
And thought of him I love.

\stropha %    2

O powerful western fallen star!
O shades of night---O moody, tearful night!
O great star disappear'd---O the black murk that hides the star!
O cruel hands that hold me powerless---O helpless soul of me!
O harsh surrounding cloud that will not free my soul.


\stropha %   3

In the dooryard fronting an old farm-house near the white-\\wash'd palings,
Stands the lilac-bush tall-growing with heart-shaped leaves of rich green,
With many a pointed blossom rising delicate, with the perfume strong I love,
With every leaf a miracle---and from this bush in the dooryard,
With delicate-color'd blossoms and heart-shaped leaves of rich green,
A sprig with its flower I break.

\stropha %    4

In the swamp in secluded recesses,
A shy and hidden bird is warbling a song.
     \vacua
Solitary the thrush,
The hermit withdrawn to himself, avoiding the settlements,
Sings by himself a song.
     \vacua
Song of the bleeding throat,
Death's outlet song of life, (for well dear brother I know,
If thou wast not granted to sing thou wouldist surely die.)

\stropha %    5

Over the breast of the spring, the land, amid cities,
Amid lanes and through old woods, where lately the violets peep'd %
    from the ground, spotting the gray debris,
Amid the grass in the fields each side of the lanes, passing the %
    endless grass,
Passing the yellow-spear'd wheat, every grain from its shroud in the %
    dark-brown fields uprisen,
Passing the apple-tree blows of white and pink in the orchards,
Carrying a corpse to where it shall rest in the grave,
Night and day journeys a coffin.

\stropha %    6

Coffin that passes through lanes and streets,
Through day and night with the great cloud darkening the land,
With the pomp of the inloop'd flags with the cities draped in black,
With the show of the States themselves as of crape-veil'd women standing,
With processions long and winding and the flambeaus of the night,
With the countless torches lit, with the silent sea of faces and the %
    unbared heads,
With the waiting depot, the arriving coffin, and the sombre faces,
With dirges through the night, with the thousand voices rising strong %
    and solemn,
With all the mournful voices of the dirges pour'd around the coffin,
The dim-lit churches and the shuddering organs---where amid these %
    you journey,
With the tolling tolling bells' perpetual clang,
Here, coffin that slowly passes,
I give you my sprig of lilac.

\stropha %    7

(Nor for you, for one alone,
Blossoms and branches green to coffins all I bring,
For fresh as the morning, thus would I chant a song for you \\O sane %
    and sacred death.
     \vacua
All over bouquets of roses,
O death, I cover you over with roses and early lilies,
But mostly and now the lilac that blooms the first,
Copious I break, I break the sprigs from the bushes,
With loaded arms I come, pouring for you,
For you and the coffins all of you O death.)

\stropha %    8

O western orb sailing the heaven,
Now I know what you must have meant as a month since I walk'd,
As I walk'd in silence the transparent shadowy night,
As I saw you had something to tell as you bent to me night after night,
As you droop'd from the sky low down as if to my side, (while the %
    other stars all look'd on,)
As we wander'd together the solemn night, (for something I know not %
    what kept me from sleep,)
As the night advanced, and I saw on the rim of the west how full you %
    were of woe,
As I stood on the rising ground in the breeze in the cool transparent night,
As I watch'd where you pass'd and was lost in the netherward black %
    of the night,
As my soul in its trouble dissatisfied sank, as where you sad orb,
Concluded, dropt in the night, and was gone.

\stropha %    9

Sing on there in the swamp,
O singer bashful and tender, I hear your notes, I hear your call,
I hear, I come presently, I understand you,
But a moment I linger, for the lustrous star has detain'd me,
The star my departing comrade holds and detains me.

\stropha %    10

O how shall I warble myself for the dead one there I loved?
And how shall I deck my song for the large sweet soul that has gone?
And what shall my perfume be for the grave of him I love?
     \vacua
Sea-winds blown from east and west,
Blown from the Eastern sea and blown from the Western sea, till %
    there on the prairies meeting,
These and with these and the breath of my chant,
I'll perfume the grave of him I love.

\stropha %    11

O what shall I hang on the chamber walls?
And what shall the pictures be that I hang on the walls,
To adorn the burial-house of him I love?
Pictures of growing spring and farms and homes,
With the Fourth-month eve at sundown, and the gray smoke lucid and bright,
With floods of the yellow gold of the gorgeous, indolent, sinking %
    sun, burning, expanding the air,
With the fresh sweet herbage under foot, and the pale green leaves %
    of the trees prolific,
In the distance the flowing glaze, the breast of the river, with a %
    wind-dapple here and there,
With ranging hills on the banks, with many a line against the sky, %
    and shadows,
And the city at hand with dwellings so dense, and stacks of chimneys,
And all the scenes of life and the workshops, and the work\-men %
    homeward returning.

\stropha %    12

Lo, body and soul---this land,
My own Manhattan with spires, and the sparkling and hurrying tides, %
    and the ships,
The varied and ample land, the South and the North in the light, %
    Ohio's shores and flashing Missouri,
And ever the far-spreading prairies cover'd with grass and corn.
     \vacua
Lo, the most excellent sun so calm and haughty,
The violet and purple morn with just-felt breezes,
The gentle soft-born measureless light,
The miracle spreading bathing all, the fulfill'd noon,
The coming eve delicious, the welcome night and the stars,
Over my cities shining all, enveloping man and land.

\stropha %    13

Sing on, sing on you gray-brown bird,
Sing from the swamps, the recesses, pour your chant from the bushes,
Limitless out of the dusk, out of the cedars and pines.
     \vacua
Sing on dearest brother, warble your reedy song,
Loud human song, with voice of uttermost woe.
     \vacua
O liquid and free and tender!
O wild and loose to my soul---O wondrous singer!
You only I hear---yet the star holds me, (but will soon depart,)
Yet the lilac with mastering odor holds me.

\stropha %    14

Now while I sat in the day and look'd forth,
In the close of the day with its light and the fields of spring, and %
    the farmers preparing their crops,
In the large unconscious scenery of my land with its lakes and forests,
In the heavenly aerial beauty, (after the perturb'd winds and the storms,)
Under the arching heavens of the afternoon swift passing, and the %
    voices of children and women,
The many-moving sea-tides, and I saw the ships how they sail'd,
And the summer approaching with richness, and the fields all busy %
    with labor,
And the infinite separate houses, how they all went on, each with %
    its meals and minutia of daily usages,
And the streets how their throbbings throbb'd, and the cities pent--- %
    lo, then and there,
Falling upon them all and among them all, enveloping me with the rest,
Appear'd the cloud, appear'd the long black trail,
And I knew death, its thought, and the sacred knowledge of death.
     \vacua
Then with the knowledge of death as walking one side of me,
And the thought of death close-walking the other side of me,
And I in the middle as with companions, and as holding the hands of %
    companions,
I fled forth to the hiding receiving night that talks not,
Down to the shores of the water, the path by the swamp in the dimness,
To the solemn shadowy cedars and ghostly pines so still.
     \vacua
And the singer so shy to the rest receiv'd me,
The gray-brown bird I know receiv'd us comrades three,
And he sang the carol of death, and a verse for him I love.
     \vacua
From deep secluded recesses,
From the fragrant cedars and the ghostly pines so still,
Came the carol of the bird.
     \vacua
And the charm of the carol rapt me,
As I held as if by their hands my comrades in the night,
And the voice of my spirit tallied the song of the bird.
     \vacua
{\Facies* \textus {\itshape}
Come lovely and soothing death,
Undulate round the world, serenely arriving, arriving,
In the day, in the night, to all, to each,
Sooner or later delicate death.
     \vacua
Prais'd be the fathomless universe,
For life and joy, and for objects and knowledge curious,
And for love, sweet love---but praise! praise! praise!
For the sure-enwinding arms of cool-enfolding death.
     \vacua
Dark mother always gliding near with soft feet,
Have none chanted for thee a chant of fullest welcome?
Then I chant it for thee, I glorify thee above all,
I bring thee a song that when thou must indeed come, come unfalteringly.
     \vacua
Approach strong deliveress,
When it is so, when thou hast taken them I joyously sing the dead,
Lost in the loving floating ocean of thee,
Laved in the flood of thy bliss O death.
     \vacua
From me to thee glad serenades,
Dances for thee I propose saluting thee, adornments and feastings for thee,
And the sights of the open landscape and the high-spread shy are fitting,
And life and the fields, and the huge and thoughtful night.
     \vacua
The night in silence under many a star,
The ocean shore and the husky whispering wave whose voice I know,
And the soul turning to thee O vast and well-veil'd death,
And the body gratefully nestling close to thee.
     \vacua
Over the tree-tops I float thee a song,
Over the rising and sinking waves, over the myriad fields and the %
    prairies wide,
Over the dense-pack'd cities all and the teeming wharves and ways,
I float this carol with joy, with joy to thee O death.
}
\stropha %    15

To the tally of my soul,
Loud and strong kept up the gray-brown bird,
With pure deliberate notes spreading filling the night.
     \vacua
Loud in the pines and cedars dim,
Clear in the freshness moist and the swamp-perfume,
And I with my comrades there in the night.
     \vacua
While my sight that was bound in my eyes unclosed,
As to long panoramas of visions.
     \vacua
And I saw askant the armies,
I saw as in noiseless dreams hundreds of battle-flags,
Borne through the smoke of the battles and pierc'd with missiles I saw them,
And carried hither and yon through the smoke, and torn and bloody,
And at last but a few shreds left on the staffs, (and all in silence,)
And the staffs all splinter'd and broken.
     \vacua
I saw battle-corpses, myriads of them,
And the white skeletons of young men, I saw them,
I saw the debris and debris of all the slain soldiers of the war,
But I saw they were not as was thought,
They themselves were fully at rest, they suffer'd not,
The living remain'd and suffer'd, the mother suffer'd,
And the wife and the child and the musing comrade suffer'd,
And the armies that remain'd suffer'd.

\stropha %    16

Passing the visions, passing the night,
Passing, unloosing the hold of my comrades' hands,
Passing the song of the hermit bird and the tallying song of my soul,
Victorious song, death's outlet song, yet varying ever-altering song,
As low and wailing, yet clear the notes, rising and falling, %
    flooding the night,
Sadly sinking and fainting, as warning and warning, and yet again %
    bursting with joy,
Covering the earth and filling the spread of the heaven,
As that powerful psalm in the night I heard from recesses,
Passing, I leave thee lilac with heart-shaped leaves,
I leave thee there in the door-yard, blooming, returning with spring.
     \vacua
I cease from my song for thee,
From my gaze on thee in the west, fronting the west, communing with thee,
O comrade lustrous with silver face in the night.
     \vacua
Yet each to keep and all, retrievements out of the night,
The song, the wondrous chant of the gray-brown bird,
And the tallying chant, the echo arous'd in my soul,
With the lustrous and drooping star with the countenance full of woe,
With the holders holding my hand nearing the call of the bird,
Comrades mine and I in the midst, and their memory ever to keep, for %
    the dead I loved so well,
For the sweetest, wisest soul of all my days and lands---and this for %
    his dear sake,
Lilac and star and bird twined with the chant of my soul,
There in the fragrant pines and the cedars dusk and dim.

\Facies \strophae {}
\stropha

\end{poem}

\begin{poem}{O Captain! My Captain!}

  \Forma \strophae {0{\penalty 1000}000234{\penalty 1000}5}
  \SpatiumInfra    {1ex plus .5ex \penalty -50}

O Captain! my Captain! our fearful trip is done,
The ship has weather'd every rack, the prize we sought is won,
The port is near, the bells I hear, the people all exulting,
While follow eyes the steady keel, the vessel grim and daring;
    But O heart! heart! heart!
     O the bleeding drops of red,
       Where on the deck my Captain lies,
         Fallen cold and dead.

O Captain! my Captain! rise up and hear the bells;
Rise up---for you the flag is flung---for you the bugle trills,
For you bouquets and ribbon'd wreaths---for you the shores a-crowding,
For you they call, the swaying mass, their eager faces turning;
    Here Captain! dear father!
     This arm beneath your head!
       It is some dream that on the deck,
         You've fallen cold and dead.

My Captain does not answer, his lips are pale and still,
My father does not feel my arm, he has no pulse nor will,
The ship is anchor'd safe and sound, its voyage closed and done,
From fearful trip the victor ship comes in with object won;
     Exult O shores, and ring O bells!
       But I with mournful tread,
         Walk the deck my Captain lies,
           Fallen cold and dead.

\end{poem}


\begin{poem}{Hush'd Be the Camps To-Day\\
            {\RelSize{-2}\textmd{[May 4, 1865]}}}

Hush'd be the camps to-day,
And soldiers let us drape our war-worn weapons,
And each with musing soul retire to celebrate,
Our dear commander's death.
     \vacua
No more for him life's stormy conflicts,
Nor victory, nor defeat---no more time's dark events,
Charging like ceaseless clouds across the sky.
But sing poet in our name,
     \vacua
Sing of the love we bore him---because you, dweller in camps, know it truly.
     \vacua
As they invault the coffin there,
Sing---as they close the doors of earth upon him---one verse,
For the heavy hearts of soldiers.

\end{poem}


\begin{poem}{This Dust Was Once the Man}

This dust was once the man,
Gentle, plain, just and resolute, under whose cautious hand,
Against the foulest crime in history known in any land or age,
Was saved the Union of these States.

\end{poem}

\begin{poem}{By Blue Ontario's Shore}

  \Modus  \excessus {\justified[850]}
  \Facies \strophae {\Snumber}

\stropha %    1

By blue Ontario's shore,
As I mused of these warlike days and of peace return'd, and the %
    dead that return no more,
A Phantom gigantic superb, with stern visage accosted me,
{\Facies* \textus {\itshape}
Chant me the poem, \textup{it said}, that comes from the soul of Ameri\-ca, %
    chant me the carol of victory,
And strike up the marches of Libertad, marches more powerful yet,
And sing me before you go the song of the throes of Democracy.
}
     \vacua
(Democracy, the destin'd conqueror, yet treacherous lip-smiles everywhere,
And death and infidelity at every step.)

\stropha %    2

A Nation announcing itself,
I myself make the only growth by which I can be appreciated,
I reject none, accept all, then reproduce all in my own forms.
     \vacua
A breed whose proof is in time and deeds,
What we are we are, nativity is answer enough to objections,
We wield ourselves as a weapon is wielded,
We are powerful and tremendous in ourselves,
We are executive in ourselves, we are sufficient in the variety of %
    ourselves,
We are the most beautiful to ourselves and in ourselves,
We stand self-pois'd in the middle, branching thence over the world,
From Missouri, Nebraska, or Kansas, laughing attacks to scorn.
     \vacua
Nothing is sinful to us outside of ourselves,
Whatever appears, whatever does not appear, we are beautiful or %
    sinful in ourselves only.
     \vacua
(O Mother---O Sisters dear!
If we are lost, no victor else has destroy'd us,
It is by ourselves we go down to eternal night.)

\stropha %    3

Have you thought there could be but a single supreme?
There can be any number of supremes---one does not countervail %
    another any more than one eyesight countervails another, or %
    one life countervails another.
     \vacua
All is eligible to all,
All is for individuals, all is for you,
No condition is prohibited, not God's or any.
     \vacua
All comes by the body, only health puts you rapport with the universe.
     \vacua
Produce great Persons, the rest follows.

\stropha %    4

Piety and conformity to them that like,
Peace, obesity, allegiance, to them that like,
I am he who tauntingly compels men, women, nations,
Crying, Leap from your seats and contend for your lives!
     \vacua
I am he who walks the States with a barb'd tongue, questioning every %
    one I meet,
Who are you that wanted only to be told what you knew before?
Who are you that wanted only a book to join you in your nonsense?
     \vacua
(With pangs and cries as thine own O bearer of many children,
These clamors wild to a race of pride I give.)
     \vacua
O lands, would you be freer than all that has ever been before?
If you would be freer than all that has been before, come listen to me.
     \vacua
Fear grace, elegance, civilization, delicatesse,
Fear the mellow sweet, the sucking of honey---juice,
Beware the advancing mortal ripening of Nature,
Beware what precedes the decay of the ruggedness of states and men.

\stropha %    5

Ages, precedents, have long been accumulating undirected materials,
America brings builders, and brings its own styles.
     \vacua
The immortal poets of Asia and Europe have done their work and %
    pass'd to other spheres,
A work remains, the work of surpassing all they have done.
     \vacua
America, curious toward foreign characters, stands by its own at all %
    hazards,
Stands removed, spacious, composite, sound, initiates the true use %
    of precedents,
Does not repel them or the past or what they have produced under %
    their forms,
Takes the lesson with calmness, perceives the corpse slowly borne %
    from the house,
Perceives that it waits a little while in the door, that it was %
    fittest for its days,
That its life has descended to the stalwart and well-shaped heir who %
    approaches,
And that he shall be fittest for his days.
     \vacua
Any period one nation must lead,
One land must be the promise and reliance of the future.
     \vacua
These States are the amplest poem,
Here is not merely a nation but a teeming Nation of nations,
Here the doings of men correspond with the broadcast doings of the %
    day and night,
Here is what moves in magnificent masses careless of particulars,
Here are the roughs, beards, friendliness, combativeness, the soul loves,
Here the flowing trains, here the crowds, equality, diversity, the %
    soul loves.

\stropha %    6

Land of lands and bards to corroborate!
Of them standing among them, one lifts to the light a west-bred face,
To him the hereditary countenance bequeath'd both mother's and father's,
His first parts substances, earth, water, animals, trees,
Built of the common stock, having room for far and near,
Used to dispense with other lands, incarnating this land,
Attracting it body and soul to himself, hanging on its neck with %
    incomparable love,
Plunging his seminal muscle into its merits and demerits,
Making its cities, beginnings, events, diversities, wars, vocal in him,
Making its rivers, lakes, bays, embouchure in him,
Mississippi with yearly freshets and changing chutes, Columbia, %
    Niagara, Hudson, spending themselves lovingly in him,
If the Atlantic coast stretch or the Pacific coast stretch, he %
    stretching with them North or South,
Spanning between them East and West, and touching whatever is between them,
Growths growing from him to offset the growths of pine, ce\-dar, hemlock, %
    live-oak, locust, chestnut, hickory, cottonwood, orange, magnolia,
Tangles as tangled in him as any canebrake or swamp,
He likening sides and peaks of mountains, forests coated with %
    northern transparent ice,
Off him pasturage sweet and natural as savanna, upland, prai\-rie,
Through him flights, whirls, screams, answering those of the %
    fish-hawk, mocking-bird, night-heron, and eagle,
His spirit surrounding his country's spirit, unclosed to good and evil,
Surrounding the essences of real things, old times and present times,
Surrounding just found shores, islands, tribes of red aborigines,
Weather-beaten vessels, landings, settlements, embryo stature and muscle,
The haughty defiance of the Year One, war, peace, the formation of %
    the Constitution,
The separate States, the simple elastic scheme, the immigrants,
The Union always swarming with blatherers and always sure and impregnable,
The unsurvey'd interior, log-houses, clearings, wild animals, hunters, trappers,
Surrounding the multiform agriculture, mines, temperature, the %
    gestation of new States,
Congress convening every Twelfth-month, the members duly coming %
    up from the uttermost parts,
Surrounding the noble character of mechanics and farmers, especially %
    the young men,
Responding their manners, speech, dress, friendships, the gait they %
    have of persons who never knew how it felt to stand in the %
    presence of superiors,
The freshness and candor of their physiognomy, the copiousness and %
    decision of their phrenology,
The picturesque looseness of their carriage, their fierceness when wrong'd,
The fluency of their speech, their delight in music, their cu\-ri\-os\-i\-ty, %
    good temper and open-handedness, the whole composite make,
The prevailing ardor and enterprise, the large amativeness,
The perfect equality of the female with the male, the fluid movement %
    of the population,
The superior marine, free commerce, fisheries, whaling, gold-\\digging,
Wharf-hemm'd cities, railroad and steamboat lines intersecting all points,
Factories, mercantile life, labor-saving machinery, the Northeast, %
    Northwest, Southwest,
Manhattan firemen, the Yankee swap, southern plantation life,
Slavery---the murderous, treacherous conspiracy to raise it\\ upon the %
    ruins of all the rest,
On and on to the grapple with it---Assassin! then your life or ours %
    be the stake, and respite no more.

\stropha %    7

(Lo, high toward heaven, this day,
Libertad, from the conqueress' field return'd,
I mark the new aureola around your head,
No more of soft astral, but dazzling and fierce,
With war's flames and the lambent lightnings playing,
And your port immovable where you stand,
With still the inextinguishable glance and the clinch'd and lifted fist,
And your foot on the neck of the menacing one, the scorner utterly %
    crush'd beneath you,
The menacing arrogant one that strode and advanced with his %
    senseless scorn, bearing the murderous knife,
The wide-swelling one, the braggart that would yesterday do so much,
To-day a carrion dead and damn'd, the despised of all the earth,
An offal rank, to the dunghill maggots spurn'd.)

\stropha %    8

Others take finish, but the Republic is ever constructive and ever %
    keeps vista,
Others adorn the past, but you O days of the present, I adorn you,
O days of the future I believe in you---I isolate myself for your sake,
O America because you build for mankind I build for you,
O well-beloved stone-cutters, I lead them who plan with decision %
    and science,
Lead the present with friendly hand toward the future.
(Bravas to all impulses sending sane children to the next age!
But damn that which spends itself with no thought of the stain, %
    pains, dismay, feebleness, it is bequeathing.)

\stropha %    9

I listened to the Phantom by Ontario's shore,
I heard the voice arising demanding bards,
By them all native and grand, by them alone can these States be %
    fused into the compact organism of a Nation.
     \vacua
To hold men together by paper and seal or by compulsion is no account,
That only holds men together which aggregates all in a living principle, %
    as the hold of the limbs of the body or the fibres of plants.
     \vacua
Of all races and eras these States with veins full of poetical stuff most %
    need poets, and are to have the greatest, and use them the greatest,
Their Presidents shall not be their common referee so much as their %
    poets shall.
     \vacua
(Soul of love and tongue of fire!
Eye to pierce the deepest deeps and sweep the world!
Ah Mother, prolific and full in all besides, yet how long barren, barren?)

\stropha %    10

Of these States the poet is the equable man,
Not in him but off from him things are grotesque, eccentric, fail of %
    their full returns,
Nothing out of its place is good, nothing in its place is bad,
He bestows on every object or quality its fit proportion, neither %
    more nor less,
He is the arbiter of the diverse, he is the key,
He is the equalizer of his age and land,
He supplies what wants supplying, he checks what wants checking,
In peace out of him speaks the spirit of peace, large, rich, %
    thrifty, building populous towns, encouraging agriculture, arts, %
    commerce, lighting the study of man, the soul, health, %
    immortality, government,
In war he is the best backer of the war, he fetches artillery as %
    good as the engineer's, he can make every word he speaks draw blood,
The years straying toward infidelity he withholds by his steady faith,
He is no arguer, he is judgment, (Nature accepts him absolutely,)
He judges not as the judge judges but as the sun failing round %
    helpless thing,
As he sees the farthest he has the most faith,
His thoughts are the hymns of the praise of things,
In the dispute on God and eternity he is silent,
He sees eternity less like a play with a prologue and denouement,
He sees eternity in men and women, he does not see men and women %
    as dreams or dots.
     \vacua
For the great Idea, the idea of perfect and free individuals,
For that, the bard walks in advance, leader of leaders,
The attitude of him cheers up slaves and horrifies foreign despots.
     \vacua
Without extinction is Liberty, without retrograde is Equality,
They live in the feelings of young men and the best women,
(Not for nothing have the indomitable heads of the earth been always %
    ready to fall for Liberty.)


\stropha %    11

For the great Idea,
That, O my brethren, that is the mission of poets.
     \vacua
Songs of stern defiance ever ready,
Songs of the rapid arming and the march,
The flag of peace quick-folded, and instead the flag we know,
Warlike flag of the great Idea.
     \vacua
(Angry cloth I saw there leaping!
I stand again in leaden rain your flapping folds saluting,
I sing you over all, flying beckoning through the fight---O the %
    hard-contested fight!
The cannons ope their rosy-flashing muzzles---the hurtled balls scream,
The battle-front forms amid the smoke---the volleys pour incessant %
    from the line,
Hark, the ringing word Charge!---now the tussle and the furious %
    maddening yells,
Now the corpses tumble curl'd upon the ground,
Cold, cold in death, for precious life of you,
Angry cloth I saw there leaping.)

\stropha %    12

Are you he who would assume a place to teach or be a poet here in %
    the States?
The place is august, the terms obdurate.
     \vacua
Who would assume to teach here may well prepare himself body and mind,
He may well survey, ponder, arm, fortify, harden, make lithe himself,
He shall surely be question'd beforehand by me with many and stern questions.
     \vacua
Who are you indeed who would talk or sing to America?
Have you studied out the land, its idioms and men?
Have you learn'd the physiology, phrenology, politics, ge\-og\-ra\-phy, %
    pride, freedom, friendship of the land? its sub\-stra\-tums and objects?
Have you consider'd the organic compact of the first day of the %
    first year of Independence, sign'd by the Commissioners, ratified %
    by the States, and read by Washington at the head of the army?
Have you possess'd yourself of the Federal Constitution?
Do you see who have left all feudal processes and poems behind them, %
    and assumed the poems and processes of Democracy?
Are you faithful to things? do you teach what the land and sea, the %
    bodies of men, womanhood, amativeness, heroic angers, teach?
Have you sped through fleeting customs, popularities?
Can you hold your hand against all seductions, follies, whirls, %
    fierce contentions? are you very strong? are you really of the %
    whole People?
Are you not of some coterie? some school or mere religion?
Are you done with reviews and criticisms of life? animating now to %
    life itself?
Have you vivified yourself from the maternity of these States?
Have you too the old ever-fresh forbearance and impartiality?
Do you hold the like love for those hardening to maturity? for the %
    last-born? little and big? and for the errant?
     \Vacua
What is this you bring my America?
Is it uniform with my country?
Is it not something that has been better told or done before?
Have you not imported this or the spirit of it in some ship?
Is it not a mere tale? a rhyme? a prettiness?---Is the good old cause in it?
Has it not dangled long at the heels of the poets, politicians, %
    literats, of enemies' lands?
Does it not assume that what is notoriously gone is still here?
Does it answer universal needs? will it improve manners?
Does it sound with trumpet-voice the proud victory of the Union in %
    that secession war?
Can your performance face the open fields and the seaside?
Will it absorb into me as I absorb food, air, to appear again in my %
    strength, gait, face?
Have real employments contributed to it? original makers, not mere %
    amanuenses?
Does it meet modern discoveries, calibres, facts, face to face?
What does it mean to American persons, progresses, cities? Chicago, %
    Kanada, Arkansas?
Does it see behind the apparent cus\-to\-dians the real cus\-to\-dians %
    standing, menacing, silent, the mechanics, Man\-hat\-tan\-ese, Western %
    men, Southerners, significant alike in their apathy, and in the %
    promptness of their love?
Does it see what finally befalls, and has always finally be\-fal\-len, %
    each temporizer, patcher, outsider, partialist, a\-larm\-ist, %
    infidel, who has ever ask'd any thing of America?
What mocking and scornful negligence?
The track strew'd with the dust of skeletons,
By the roadside others disdainfully toss'd.

\stropha %    13

Rhymes and rhymers pass away, poems distill'd from poems pass away,
The swarms of reflectors and the polite pass, and leave ashes,
Admirers, importers, obedient persons, make but the soil of literature,
America justifies itself, give it time, no disguise can deceive it %
    or conceal from it, it is impassive enough,
Only toward the likes of itself will it advance to meet them,
If its poets appear it will in due time advance to meet them, there %
    is no fear of mistake,
(The proof of a poet shall be sternly deferr'd till his country %
    absorbs him as affectionately as he has absorb'd it.)
     \vacua
He masters whose spirit masters, he tastes sweetest who results %
    sweetest in the long run,
The blood of the brawn beloved of time is unconstraint;
In the need of songs, philosophy, an appropriate native grand-opera, %
    shipcraft, any craft,
He or she is greatest who contributes the greatest original %
    practical example.
     \vacua
Already a nonchalant breed, silently emerging, appears on the streets,
People's lips salute only doers, lovers, satisfiers, positive knowers,
There will shortly be no more priests, I say their work is done,
Death is without emergencies here, but life is perpetual emergencies here,
Are your body, days, manners, superb? after death you shall be superb,
Justice, health, self-esteem, clear the way with irresistible po-\\wer;
How dare you place any thing before a man?

\stropha %    14

Fall behind me States!
A man before all---myself, typical, before all.
     \vacua
Give me the pay I have served for,
Give me to sing the songs of the great Idea, take all the rest,
I have loved the earth, sun, animals, I have despised riches,
I have given aims to every one that ask'd, stood up for the stupid %
    and crazy, devoted my income and labor to others,
Hated tyrants, argued not concerning God, had patience and indulgence % 
    toward the people, taken off my hat to nothing known or unknown,
\DriveOut Gone freely with powerful uneducated persons and with the young, %
    and with the mothers of families,
Read these leaves to myself in the open air, tried them by trees, %
    stars, rivers,
Dismiss'd whatever insulted my own soul or defiled my body,
Claim'd nothing to myself which I have not carefully claim'd for %
    others on the same terms,
Sped to the camps, and comrades found and accepted from every State,
(Upon this breast has many a dying soldier lean'd to breathe his last,
This arm, this hand, this voice, have nourish'd, rais'd, restored,
To life recalling many a prostrate form;)
I am willing to wait to be understood by the growth of the taste of myself,
Rejecting none, permitting all.
     \vacua
(Say O Mother, have I not to your thought been faithful?
Have I not through life kept you and yours before me?)

\stropha %    15

I swear I begin to see the meaning of these things,
It is not the earth, it is not America who is so great,
It is I who am great or to be great, it is You up there, or any one,
It is to walk rapidly through civilizations, governments,theories,
Through poems, pageants, shows, to form individuals.
     \vacua
Underneath all, individuals,
I swear nothing is good to me now that ignores individuals,
The American compact is altogether with individuals,
The only government is that which makes minute of individuals,
The whole theory of the universe is directed unerringly to one %
    single individual---namely to You.
     \vacua
(Mother! with subtle sense severe, with the naked sword in your hand,
I saw you at last refuse to treat but directly with individuals.)

\stropha %    16

Underneath all, Nativity,
I swear I will stand by my own nativity, pious or impious so\\be it;
I swear I am charm'd with nothing except nativity,
Men, women, cities, nations, are only beautiful from nativity.
     \vacua
Underneath all is the Expression of love for men and women,
(I swear I have seen enough of mean and impotent modes of expressing %
    love for men and women,
After this day I take my own modes of expressing love for men and %
    women.) in myself,
     \vacua
I swear I will have each quality of my race in myself,
(Talk as you like, he only suits these States whose manners favor %
    the audacity and sublime turbulence of the States.)
     \vacua
Underneath the lessons of things, spirits, Nature, governments, %
    ownerships, I swear I perceive other lessons,
Underneath all to me is myself, to you yourself, (the same %
    monotonous old song.)

\stropha %    17

O I see flashing that this America is only you and me,
Its power, weapons, testimony, are you and me,
Its crimes, lies, thefts, defections, are you and me,
Its Congress is you and me, the officers, capitols, armies, ships, %
    are you and me,
Its endless gestations of new States are you and me,
The war, (that war so bloody and grim, the war I will henceforth %
    forget), was you and me,
Natural and artificial are you and me,
Freedom, language, poems, employments, are you and me,
Past, present, future, are you and me.
     \vacua
I dare not shirk any part of myself,
Not any part of America good or bad,
Not to build for that which builds for mankind,
Not to balance ranks, complexions, creeds, and the sexes,
Not to justify science nor the march of equality,
Nor to feed the arrogant blood of the brawn belov'd of time.
     \vacua
I am for those that have never been master'd,
For men and women whose tempers have never been master'd,
For those whom laws, theories, conventions, can never master.
     \vacua
I am for those who walk abreast with the whole earth,
Who inaugurate one to inaugurate all.
     \vacua
I will not be outfaced by irrational things,
I will penetrate what it is in them that is sarcastic upon me,
I will make cities and civilizations defer to me,
This is what I have learnt from America---it is the amount, and it I %
    teach again.
     \vacua
(Democracy, while weapons were everywhere aim'd at your breast,
I saw you serenely give birth to immortal children, saw in dreams %
    your dilating form,
Saw you with spreading mantle covering the world.)

\stropha %    18

I will confront these shows of the day and night,
I will know if I am to be less than they,
I will see if I am not as majestic as they,
I will see if I am not as subtle and real as they,
I will see if I am to be less generous than they,
I will see if I have no meaning, while the houses and ships have meaning,
I will see if the fishes and birds are to be enough for themselves, %
    and I am not to be enough for myself.
     \vacua
I match my spirit against yours you orbs, growths, mountains, brutes,
Copious as you are I absorb you all in myself, and become the master myself,
America isolated yet embodying all, what is it finally except myself?
These States, what are they except myself?
     \vacua
I know now why the earth is gross, tantalizing, wicked, it is for my sake,
I take you specially to be mine, you terrible, rude forms.
     \vacua

(Mother, bend down, bend close to me your face,
I know not what these plots and wars and deferments are for,
I know not fruition's success, but I know that through war and crime %
    your work goes on, and must yet go on.)

\stropha %    19

Thus by blue Ontario's shore,
While the winds fann'd me and the waves came trooping toward me,
I thrill'd with the power's pulsations, and the charm of my theme %
    was upon me,
Till the tissues that held me parted their ties upon me.
     \vacua
And I saw the free souls of poets,
The loftiest bards of past ages strode before me,
Strange large men, long unwaked, undisclosed, were disclosed to me.

\stropha %    20

O my rapt verse, my call, mock me not!
Not for the bards of the past, not to invoke them have I launch'd %
    you forth,
Not to call even those lofty bards here by Ontario's shores,
Have I sung so capricious and loud my savage song.
     \vacua
Bards for my own land only I invoke,
(For the war the war is over, the field is clear'd,)
Till they strike up marches henceforth triumphant and onward,
To cheer O Mother your boundless expectant soul.
     \vacua
Bards of the great Idea! bards of the peaceful inventions! (for the %
    war, the war is over!)
Yet bards of latent armies, a million soldiers waiting ever-ready,
Bards with songs as from burning coals or the lightning's fork'd stripes!
Ample Ohio's, Kanada's bards---bards of California! inland bards\\---%
    bards of the war!
You by my charm I invoke.

\Facies \strophae {}

\end{poem}


\begin{poem}{Reversals}

Let that which stood in front go behind,
Let that which was behind advance to the front,
Let bigots, fools, unclean persons, offer new propositions,
Let the old propositions be postponed,
Let a man seek pleasure everywhere except in himself,
Let a woman seek happiness everywhere except in herself.


\end{poem}

\end{document}
