\documentclass{book}
\usepackage[pagestyles,outermarks,clearempty]{titlesec}[2005/01/22 v2.6]
\usepackage{titletoc}[2005/01/22 v1.5]
\usepackage[repeat]{poetry}
\usepackage{drama}
\usepackage{example}

\TextHeight {5.5in}
\TextWidth  {4.5in}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


  \titleformat  {\chapter}[block]
                {\RelSize{1}\centering}
                {}
                {0pt}
                {\LETTERspace}

  \renewcommand {\contentsname}{Contents}

  \titlecontents {section}[0pt]
                 {\addvspace {1\leading}}
                 {\textsc}
                 {\textsc}
                 {\hfill \OldStyleNums{\thecontentspage}}
                 [\addvspace{.33ex}]

 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newpagestyle {MainMatterPage} {

  \sethead   [\oldstylenums{\thepage}]
             [\textsc{the metaphysical poets}]
             []
             {}
             {\scshape \toptitlemarks \MakeLowercase{\sectiontitle}}
             {\oldstylenums{\thepage}}

  \Capita    {section}{subsection}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\Novus \titulus \name
\Facies         {\newpage\thispagestyle{empty}\LETTERspace{#1}} 
\SpatiumSupra   {2\leading plus .5\leading \penalty 10000}
\Caput          {\section \indexes[a]}

\Novus \titulus \dates
\Facies         {\RelSize{+1}(\oldstylenums{#1})} 
\SpatiumSupra   {.25\leading plus .125\leading}
\SpatiumInfra   {1\leading plus .25\leading minus .25\leading}

\newcommand {\poet}[2]{\name{#1}\dates{#2}}

\Novus \titulus  \poemtitle
\Facies           {\RelSize{+1}\itshape#1}
\SpatiumSupra     {1.5\leading plus .5\leading minus .5\leading
                   \penalty (.2)}
\SpatiumInfra     {.5\leading plus .25\leading minus .25\leading
                   \penalty 10000}
\Caput            {\indexes}

\SpatiumSupra \stropham {.25\leading plus .125\leading minus .125\leading 
                         \penalty -100}

\Locus \numeri {\rightmargin}
\Facies        {\RelSize{-1}\oldstylenums{#1}}

\Locus \textus   {\auto}

\Locus \excessus {\textrightedge}
\Facies          {[\,#1}
\Modus           {\aligned{right}}

\BuildIndexes [lat']{yes}

\newenvironment{poem}[1]%
  {\poemtitle{#1}%
   \Versus
   \numerus{1}%
  }
  {\endVersus}

\Novum \spatium \LastButOne
\Facies         {\penalty 10000}

\begin{document}


\ExampleTitle {The metaphysical poets}{\itshape An Anthology}
              {The Metaphysical Poets\\[.5ex] Penguin Books, 1966}

\addtocontents{toc}{\protect\thispagestyle{empty}}
\addtocontents{toc}{\protect\setcounter{tocdepth}{2}}
\pagenumbering{roman}
\tableofcontents

\newpage
\pagestyle {MainMatterPage} 
\pagenumbering{arabic}
\thispagestyle{empty}

\Indexes

\poet{John Donne\indexes{Donne, John}}{1572\textendash1631}

\begin{poem}{Song\addfirstline}   %1
\Forma \strophae {01010133 \LastButOne 0}
  Goe, and catche a falling starre,  \auto 
    Get with child a mandrake roote, 
  Tell me, where all past yeares are,        
    Or who cleft the Divels foot,        
  Teach me to heare Mermaides singing,       
    Or to keep off envies stinging,        
            And finde        
            What winde        
  Serves to advance an honest minde. 

  If thou beest borne to strange sights,       
    Things invisible to see,        
  Ride ten thousand daies and nights,        
    Till age snow white haires on thee,        
  Thou, when thou retorn'st, wilt tell mee        
  All strange wonders that befell thee,       
            And sweare        
            No where        
  Lives a woman true, and faire.        

  If thou findst one, let mee know,        
    Such a Pilgrimage were sweet;       
  Yet doe not, I would not goe,        
    Though at next doore wee might meet,        
  Though shee were true, when you met her,        
  And last, till you write your letter,        
            Yet shee       
            Will bee        
  False, ere I come, to two, or three.        
\end{poem}

\begin{poem}{The Sunne Rising}   %2
\Forma \strophae {220022110 \LastButOne 0}
      
          Busie old foole, unruly Sunne,   \NotFirstLine     
          Why dost thou thus,        
  Through windowes, and through curtaines call on us? \auto       
  Must to thy motions lovers seasons run?          
          Sawcy pedantique wretch, goe chide       
          Late schoole boyes, and sowre prentices,        
    Goe tell Court-huntsmen, that the King will ride,        
    Call countrey ants to harvest offices;        
  Love, all alike, no season knowes, nor clyme,        
  Nor houres, dayes, moneths, which are the rags of time.       
       
          Thy beames, so reverend, and strong        
          Why shouldst thou thinke?        
  I could eclipse and cloud them with a winke,        
  But that I would not lose her sight so long:        
          If her eyes have not blinded thine,       
          Looke, and to morrow late, tell mee,        
    Whether both the'India's of spice and Myne        
    Be where thou leftst them, or lie here with mee.        
  Aske for those Kings whom thou saw'st yesterday,        
  And thou shalt heare, All here in one bed lay.       

          She'is all States, and all Princes, I,        
          Nothing else is.        
  Princes doe but play us; compar'd to this,        
  All honor's mimique; All wealth alchimie.        
          Thou sunne art halfe as happy'as wee,       
          In that the world's contracted thus;        
    Thine age askes ease, and since thy duties bee        
    To warme the world, that's done in warming us.        
  Shine here to us, and thou art every where;        
  This bed thy center is, these walls, thy spheare.       
\end{poem}
    
\begin{poem}{Song\addfirstline}  %3
\Forma \strophae {0101200 \LastButOne 1}
      
  Sweetest love, I do not goe,        
    For wearinesse of thee,        
  Nor in hope the world can show    \auto       
    A fitter Love for mee;        
      But since that I       
  Must dye at last, 'tis best,        
  To use my selfe in jest        
    Thus by fain'd deaths to dye;        
       
  Yesternight the Sunne went hence,        
    And yet is here to day,       
  He hath no desire nor sense,        
    Nor halfe so short a way:        
      Then feare not mee,        
  But beleeve that I shall make        
  Speedier journeyes, since I take       
    More wings and spurres then hee.        
       
  O how feeble is mans power        
    That if good fortune fall,        
  Cannot adde another houre,        
    Nor a lost houre recall!       
      But come bad chance,        
  And wee joyne to'it our strength,        
  And wee teach it art and length,        
    It selfe o'r us to'advance.        
       
  When thou sigh'st, thou sigh'st not winde,       
    But sigh'st my soule away,        
  When thou weep'st, unkindly kinde,        
    My lifes blood doth decay.        
      It cannot bee        
  That thou lov'st mee, as thou say'st,       
  If in thine my life thou waste,        
    Thou art the best of mee.        
       
  Let not thy divining heart        
    Forethinke me any ill,        
  Destiny may take thy part,       
    And may thy feares fulfill;        
      But thinke that wee        
  Are but turn'd aside to sleepe;        
  They who one another keepe        
    Alive, ne'r parted bee.       
\end{poem}
      
    
\begin{poem}{Aire and Angels} %4
\Forma \strophae {0000101010120 \LastButOne 0}
      
 Twice or thrice had I loved thee,        
  Before I knew thy face or name;        
  So in a voice, so in a shapelesse flame,     \auto      
   affect us oft, and worship'd bee;        
    Still when, to where thou wert, I came,       
  Some lovely glorious nothing I did see.        
    But since my soule, whose child love is,        
  Takes limmes of flesh, and else could nothing doe,        
    More subtile then the parent is,        
  Love must not be, but take a body too,       
    And therefore what thou wert, and who,        
        I bid Love aske, and now        
  That it assume thy body, I allow,        
  And fixe it selfe in thy lip, eye, and brow.        
       
  Whilst thus to ballast love, I thought,       
  And so more steddily to have gone,        
  With wares which would sinke admiration,        
  I saw, I had loves pinnace overfraught,        
    Ev'ry thy haire for love to worke upon        
  Is much too much, some fitter must be sought;       
    For, nor in nothing, nor in things        
  Extreme, and scatt'ring bright, can love inhere;        
    Then as an Angell, face, and wings        
  Of aire, not pure as it, yet pure doth weare,        
    So thy love may be my loves spheare;       
        Just such disparitie        
  As is twixt Aire and Angells puritie,        
  'Twixt womens love, and mens will ever bee.        
\end{poem}
    
      
\begin{poem}{The Anniversarie} %5
\Forma \strophae {110000100 \LastButOne 0}
    All Kings, and all their favorites,        
    All glory of honors, beauties, wits,        
  The Sun it selfe, which makes times, as they passe,  \auto
  Is elder by a yeare, now, then it was        
  When thou and I first one another saw:       
  All other things, to their destruction draw,        
    Only our love hath no decay;        
  This, no to morrow hath, nor yesterday,        
  Running it never runs from us away,        
  But truly keepes his first, last, everlasting day.       
       
    Two graves must hide thine and my coarse,        
    If one might, death were no divorce.        
  Alas, as well as other Princes, wee,        
  (Who Prince enough in one another bee,)        
  Must leave at last in death, these eyes, and eares,       
  Oft fed with true oathes, and with sweet salt teares;        
    But soules where nothing dwells but love        
  (All other thoughts being inmates) then shall prove        
  This, or a love increased there above,        
  When bodies to their graves, soules from their graves remove.       
       
    And then wee shall be throughly blest,        
    But wee no more, then all the rest;        
  Here upon earth, we'are Kings, and none but wee        
  Can be such Kings, nor of such subjects bee.        
  Who is so safe as wee? where none can doe       
  Treason to us, except one of us two.        
    True and false feares let us refraine,        
  Let us love nobly, and live, and adde againe        
  Yeares and yeares unto yeares, till we attaine        
  To write threescore: this is the second of our raigne.       
\end{poem}
      
    
\begin{poem}{Twicknam garden}  %6
\Forma \strophae {01101010 \LastButOne 1}
\Locus \textus {\previous}      
      
  Blasted with sighs, and surrounded with teares,        
    Hither I come to seeke the spring, \auto       
    And at mine eyes, and at mine eares,        
  Receive such balmes, as else cure every thing;        
    But O, selfe traytor, I do bring       
  The spider love, which transubstantiates all,        
    And can convert Manna to gall,        
  And that this place may thoroughly be thought        
    True Paradise, I have the serpent brought.        
       
  'Twere wholsomer for mee, that winter did       
    Benight the glory of this place,        
    And that a grave frost did forbid        
  These trees to laugh, and mocke mee to my face;        
    But that I may not this disgrace        
  Indure, nor yet leave loving, Love let mee       
    Some senslesse peece of this place bee;        
  Make me a mandrake, so I may groane here,        
    Or a stone fountaine weeping out my yeare.        
       
  Hither with christall vyals, lovers come,        
    And take my teares, which are loves wine,       
    And try your mistresse Teares at home,        
  For all are false, that tast not just like mine;        
    Alas, hearts do not in eyes shine,        
  Nor can you more judge womans thoughts by teares,        
    Then by her shadow, what she weares.       
  O perverse sexe, where none is true but shee,        
    Who's therefore true, because her truth kills mee.        
\end{poem}
      
    
\begin{poem}{The Dreame} %7
\Forma \strophae {0020*6 \LastButOne 0}
      
  Deare love, for nothing lesse then thee        
  Would I have broke this happy dreame,        
          It was a theame        
  For reason, much too strong for phantasie,        
  Therefore thou wakd'st me wisely; yet       
  My Dreame thou brok'st not, but continued'st it,        
  Thou art so truth, that thoughts of thee suffice,        
  To make dreames truths; and fables histories;        
  Enter these armes, for since thou thoughtst it best, \auto       
  Not to dreame all my dreame, let's act the rest.       
       
  As lightning, or a Tapers light,        
  Thine eyes, and not thy noise wak'd mee;        
          Yet I thought thee        
  (For thou lovest truth) an Angell, at first sight,        
  But when I saw thou sawest my heart,       
  And knew'st my thoughts, beyond an Angels art,        
  When thou knew'st what I dreamt, when thou knew'st when        
  Excesse of joy would wake me, and cam'st then,        
  I must confesse, it could not chuse but bee        
  Prophane, to thinke thee any thing but thee.       
       
  Comming and staying show'd thee, thee,        
  But rising makes me doubt, that now,        
          Thou art not thou.        
  That love is weake, where feare's as strong as hee;        
  'Tis not all spirit, pure, and brave,       
  If mixture it of  have.        
  Perchance as torches which must ready bee,        
  Men light and put out, so thou deal'st with mee,        
  Thou cam'st to kindle, goest to come; Then I        
  Will dreame that hope againe, but else would die.       
\end{poem}
      
    
      
\begin{poem}{A Valediction: of weeping} %8
\Forma \strophae {20002200 \LastButOne 0}
      
          Let me powre forth        
  My teares before thy face, whil'st I stay here,        
  For thy face coines them, and thy stampe they beare, \auto       
  And by this Mintage they are something worth,        
          For thus they bee       
          Pregnant of thee;        
  Fruits of much griefe they are, emblemes of more,        
  When a teare falls, that thou falst which it bore,        
  So thou and I are nothing then, when on a divers shore.        
       
          On a round ball       
  A workeman that hath copies by, can lay        
  An Europe, Afrique, and an Asia,        
  And quickly make that, which was nothing,         
          So doth each teare,        
          Which thee doth weare,       
  A globe, yea world by that impression grow,        
  Till thy teares mixt with mine doe overflow        
  This world, by waters sent from thee, my heaven dissolved so.        
       
          O more then Moone,        
  Draw not up seas to drowne me in thy spheare,       
  Weepe me not dead, in thine armes, but forbeare        
  To teach the sea, what it may doe too soone;        
          Let not the winde        
          Example finde,        
  To doe me more harme, then it purposeth;       
  Since thou and I sigh one anothers breath,        
  Who e'r sighes most, is cruellest, and hasts the others death.        
\end{poem}
      
\begin{poem}{A nocturnall upon S. Lucies day,\\ Being the shortest day} %9
\Forma \strophae {00112000 \LastButOne 0}
      
  'Tis the yeares midnight, and it is the dayes,  \auto      
   who scarce seaven houres herself unmaskes,        
    The Sunne is spent, and now his flasks        
    Send forth light squibs, no constant rayes;        
      The worlds whole sap is sunke:       
  The generall balme th'hydroptique earth hath drunk,        
  Whither, as to the beds-feet, life is shrunke,        
  Dead and enterr'd; yet all these seeme to laugh,        
  Compar'd with mee, who am their Epitaph.        
       
  Study me then, you who shall lovers bee       
  At the next world, that is, at the next Spring:        
    For I am every dead thing,        
    In whom love wrought new Alchimie.        
      For his art did expresse        
  A quintessence even from nothingnesse,       
  From dull privations, and leane emptinesse:        
  He ruin'd mee, and I am re-begot        
  Of absence, darknesse, death; things which are not.        
       
  All others, from all things, draw all that's good,        
  Life, soule, forme, spirit, whence they beeing have;       
    I, by loves limbecke, am the grave        
    Of all, that's nothing. Oft a flood        
      Have wee two wept, and so        
  Drownd the whole world, us two; oft did we grow        
  To be two Chaosses, when we did show       
  Care to ought else; and often absences        
  Withdrew our soules, and made us carcasses.        
       
  But I am by her death, (which word wrongs her)        
  Of the first nothing, the Elixer grown;        
    Were I a man, that I were one,       
    I needs must know; I should preferre,        
      If I were any beast,        
  Some ends, some means; Yea plants, yea stones detest,        
  And love; All, all some properties invest;        
  If I an ordinary nothing were,       
  As shadow, a light, and body must be here.        
       
  But I am None; nor will my Sunne renew.        
  You lovers, for whose sake, the lesser Sunne        
    At this time to the Goat is runne        
    To fetch new lust, and give it you,       
      Enjoy your summer all;        
  Since shee enjoyes her long nights festivall,        
  Let mee prepare towards her, and let mee call        
  This houre her Vigill, and her Eve, since this        
  Both the yeares, and the dayes deep midnight is.       
\end{poem}      
    
\begin{poem}{A Valediction: forbidding mourning} %10
\Forma \strophae {010 \LastButOne 1}
      
  As virtuous men passe mildly away,        
    And whisper to their soules, to goe,        
  Whilst some of their sad friends doe say, \auto       
    The breath goes now, and some say, no:        
       
  So let us melt, and make no noise,       
    No teare-floods, nor sigh-tempests move,        
  T'were prophanation of our joyes        
    To tell the layetie our love.        
       
  Moving of th'earth brings harmes and feares,        
    Men reckon what it did and meant,       
  But trepidation of the spheares,        
    Though greater farre, is innocent.        
       
  Dull sublunary lovers love        
    (Whose soule is sense) cannot admit        
  Absence, because it doth remove       
    Those things which elemented it.        
       
  But we by a love, so much refin'd,        
    That our selves know not what it is,        
  Inter-assured of the mind,        
    Care lesse, eyes, lips, and hands to misse.       
       
  Our two soules therefore, which are one,        
    Though I must goe, endure not yet        
  A breach, but an expansion,        
    Like gold to ayery thinnesse beate.        
       
  If they be two, they are two so       
    As stiffe twin compasses are two,        
  Thy soule the fixt foot, makes no show        
    To move, but doth, if the'other doe.        
       
  And though it in the center sit,        
    Yet when the other far doth rome,       
  It leanes, and hearkens after it,        
    And growes erect, as that comes home.        
       
  Such wilt thou be to mee, who must        
    Like th'other foot, obliquely runne;        
  Thy firmnes makes my circle just,       
    And makes me end, where I begunne.        
\end{poem}      
      
    
\begin{poem}{The Extasie} %11
\Forma \strophae {010 \LastButOne 1}
      
  Where, like a pillow on a bed,        
    A Pregnant banke swel'd up, to rest \auto       
  The violets reclining head,        
    Sat we two, one anothers best.        
  Our hands were firmely cimented       
    With a fast balme, which thence did spring,        
  Our eye-beames twisted, and did thred        
    Our eyes, upon one double string;        
  So to'entergraft our hands, as yet        
    Was all the meanes to make us one,       
  And pictures in our eyes to get        
    Was all our propagation.        
  As 'twixt two equall Armies, Fate        
    Suspends uncertaine victorie,        
  Our soules, (which to advance their state,       
    Were gone out,) hung 'twixt her, and mee.        
  And whil'st our soules negotiate there,        
    Wee like sepulchrall statues lay;        
  All day, the same our postures were,        
    And wee said nothing, all the day.       
  If any, so by love refin'd,        
    That he soules language understood,        
  And by good love were growen all minde,        
    Within convenient distance stood,        
  He (though he knew not which soule spake,       
    Because both meant, both spake the same)        
  Might thence a new concoction take,        
    And part farre purer then he came.        
  This Extasie doth unperplex        
    (We said) and tell us what we love,       
  Wee see by this, it was not sexe,        
    Wee see, we saw not what did move:        
  But as all severall soules containe        
    Mixture of things, they know not what,        
  Love, these mixt soules, doth mixe againe,       
    And makes both one, each this and that.        
  A single violet transplant,        
    The strength, the colour, and the size,        
  (All which before was poore, and scant,)        
    Redoubles still, and multiplies.       
  When love, with one another so        
    Interinanimates two soules,        
  That abler soule, which thence doth flow,        
    Defects of lonelinesse controules.        
  Wee then, who are this new soule, know,       
    Of what we are compos'd, and made,        
  For, th'Atomies of which we grow,        
    Are soules, whom no change can invade.        
  But O alas, so long, so farre        
    Our bodies why doe wee forbeare?       
  They are ours, though they are not wee, Wee are        
    The intelligences, they the spheare.        
  We owe them thankes, because they thus,        
    Did us, to us, at first convay,        
  Yeelded their forces, sense, to us,       
    Nor are drosse to us, but allay.        
  On man heavens influence workes not so,        
    But that it first imprints the ayre,        
  Soe soule into the soule may flow,        
    Though it to body first repaire.       
  As our blood labours to beget        
    Spirits, as like soules as it can,        
  Because such fingers need to knit        
    That subtile knot, which makes us man:        
  So must pure lovers soules descend       
    T'affections, and to faculties,        
  Which sense may reach and apprehend,        
    Else a great Prince in prison lies.        
  To'our bodies turne wee then, that so        
    Weake men on love reveal'd may looke;       
  Loves mysteries in soules doe grow,        
    But yet the body is his booke.        
  And if some lover, such as wee,        
    Have heard this dialogue of one,        
  Let him still marke us, he shall see       
    Small change, when we'are to bodies gone.        
\end{poem}      
      
    
\begin{poem}{The Relique} %12
\Forma \strophae {1111202000 \LastButOne 0}
      
      When my grave is broke up againe  \auto      
      Some second ghest to entertaine,        
      (For graves have learn'd that woman-head        
      To be to more then one a Bed)        
          And he that digs it, spies       
  A bracelet of bright haire about the bone,        
          Will he not let'us alone,        
  And thinke that there a loving couple lies,        
  Who thought that this device might be some way        
  To make their soules, at the last busie day,       
  Meet at this grave, and make a little stay?        
       
          If this fall in a time, or land,        
          Where mis-devotion doth command,        
          Then, he that digges us up, will bring        
          Us, to the Bishop, and the King,       
            To make us Reliques; then        
  Thou shalt be a Mary Magdalen, and I        
            A something else thereby;        
  All women shall adore us, and some men;        
  And since at such time, miracles are sought,       
  I would have that age by this paper taught        
  What miracles wee harmelesse lovers wrought.        
       
          First, we lov'd well and faithfully,        
          Yet knew not what wee lov'd, nor why,        
          Difference of sex no more wee knew,       
          Then our Guardian Angells doe;        
            Comming and going, wee        
  Perchance might kisse, but not between those meales;        
            Our hands ne'r toucht the seales,        
  Which nature, injur'd by late law, sets free:       
  These miracles wee did; but now alas,        
  All measure, and all language, I should passe,        
  Should I tell what a miracle shee was.        
\end{poem}      


\poet{George Herbert\indexes{Herbert, George}}{1593\textendash1633}

\begin{poem}{The Collar} %13  
\Forma \strophae {03101211130212031310030130330323031 \LastButOne 3\\2em}      
      I struck the board, and cry'd, No more.        
                          I will abroad.        
      What? shall I ever sigh and pine?        
  My lines and life are free; free as the rode,        
      Loose as the winde, as large as store.       
            Shall I be still in suit?        
      Have I no harvest but a thorn        
      To let me bloud, and not restore        
  What I have lost with cordiall fruit?        
                          Sure there was wine       
        Before my sighs did drie it: there was corn        
            Before my tears did drown it.        
      Is the yeare onely lost to me?        
            Have I no bayes to crown it?        
  No flowers, no garlands gay? all blasted?       
                          All wasted?        
      Not so, my heart: but there is fruit,        
                          And thou hast hands.        
            Recover all thy sigh-blown age        
  On double pleasures: leave thy cold dispute       
  Of what is fit, and not; forsake thy cage,        
                          Thy rope of sands,        
  Which pettie thoughts have made, and made to thee        
      Good cable, to enforce and draw,        
                          And be thy law,       
      While thou didst wink and wouldst not see.        
                          Away; take heed:        
                          I will abroad.        
  Call in thy deaths head there: tie up thy fears.        
                          He that forbears       
            To suit and serve his need,        
                          Deserves his load.        
  But as I rav'd and grew more fierce and wilde        
                          At every word,        
      Me thoughts I heard one calling, \textit{Childe!}       
                And I reply'd, \textit{My Lord}.        
\end{poem}      
\begin{poem}{Aaron}   
\Forma \strophae {2101 \LastButOne 2\\2em}      
      
          Holinesse on the head,        
      Light and perfections on the breast,        
  Harmonious bells below, raising the dead        
      To leade them unto life and rest.        
          Thus are true Aarons drest.       
          Profanenesse in my head,        
      Defects and darknesse in my breast,        
  A noise of passions ringing me for dead        
      Unto a place where is no rest.        
          Poore priest thus am I drest.       
          Onely another head        
      I have, another heart and breast,        
  Another musick, making live not dead,        
      Without whom I could have no rest:        
          In him I am well drest.       
          Christ is my onely head,        
      My alone onely heart and breast,        
  My onely musick, striking me ev'n dead;        
      That to the old man I may rest,        
          And be in him new drest.       
       
          So holy in my head,        
      Perfect and light in my deare breast,        
  My doctrine tun'd by Christ, (who is not dead,        
      But lives in me while I do rest)        
          Come people; Aaron 's drest.       
      

\end{poem}   
\begin{poem}{Discipline}   
\Forma \strophae {003 \LastButOne 0}      
      
  Throw away thy rod,        
  Throw away thy wrath: \auto       
                O my God,        
  Take the gentle path.        
       
  For my hearts desire       
  Unto thine is bent:        
                I aspire        
  To a full consent.        
       
  Not a word or look        
  I affect to own,       
                But by book,        
  And thy book alone.        
       
  Though I fail, I weep:        
  Though I halt in pace,        
                Yet I creep       
  To the throne of grace.        
       
  Then let wrath remove;        
  Love will do the deed:        
                For with love        
  Stonie hearts will bleed.       
       
  Love is swift of foot;        
  Love 's a man of warre,        
                And can shoot,        
  And can hit from farre.        
       
  Who can scape his bow?       
  That which wrought on thee,        
                Brought thee low,        
  Needs must work on me.        
       
  Throw away thy rod;        
  Though man frailties hath,       
                Thou art God:        
  Throw away thy wrath.        
      
\end{poem}
\begin{poem}{Redemption}   
\Forma \strophae {0110011001101 \LastButOne 1}      
      
  Having been tenant long to a rich Lord,  \auto      
        Not thriving, I resolved to be bold,        
        And make a suit unto him, to afford        
  A new small-rented lease, and cancell th' old.        
       
  In heaven at his manour I him sought:       
        They told me there, that he was lately gone        
        About some land, which he had dearly bought        
  Long since on earth, to take possession.        
       
  I straight return'd, and knowing his great birth,        
        Sought him accordingly in great resorts;       
        In cities, theatres, gardens, parks, and courts:        
  At length I heard a ragged noise and mirth        
       
        Of theeves and murderers: there I him espied,        
        Who straight, \textit{Your suit is granted}, said, \& died.        
      
\end{poem}

\newpage
\begin{poem}{Easter-wings}   
\Forma \strophae [10]{c{.5ex}}
\Locus \textus {+0pt}
Lord, who createdst man in wealth and store, 
Though foolishly he lost the same,
Decaying more and more,
Till he became
Most poore
With thee
O let me rise
As larks, harmoniously,
And sing this day thy victories:
Then shall the fall further the flight in me.
My tender age in sorrow did beginne:
And still with sicknesses and shame
Thou didst so punish sinne,
That I became
Most thinne.
With thee
Let me combine,
And feel this day thy victorie:
For, if I imp my wing on thine,
Affliction shall advance the flight in me.
\end{poem}

\poet{Richard Crashaw\indexes{Crashaw, Richard}}{1612\textendash1649}
      
\begin{poem}{Hymn of the Nativity\\Sung as by the Shepheards}   
\Drama
\Facies\personae{\textsc{#1}}
\Locus\personae{\centre}
\SpatiumSupra\personam{1ex plus .25ex minus .25ex\penalty -100}
\SpatiumInfra\personam{.33ex plus .11ex\penalty 10000}
\persona*[1]{Chorus}
\persona*[2]{Tityrus}
\persona*[3]{Thyrsis}
\persona*[4]{Both}
\persona*[5]{Full Chorus}
\Forma \strophae {010 \LastButOne 1}      
\Locus \textus {\auto}

  \1  Come we shepheards whose blest Sight \auto
  Hath mett love's Noon in Nature's night;        
    Come lift we up our loftyer Song        
  And wake the \textit{Sun} that lyes too long.        

\Forma \strophae {01010 \LastButOne 0}      
       
    To all our world of well-stoln joy       
  He slept; and dream't of no such thing;        
    While we found out Heavn's fairer eye        
  And Kis't the Cradle of our \textit{King}.        
    Tell him He rises now too late        
  To show us ought worth looking at.       
       
    Tell him we now can show Him more        
  Then He e're show'd to mortall Sight;        
    Then he Himselfe e're saw before;        
  Which to be seen needes not His light.        
    Tell him, Tityrus, where th' hast been,       
  Tell him, Thyrsis, what th' hast seen.        

    \2 Gloomy night embrac't the Place        
  Where The Noble Infant lay.        
    The \textit{Babe} look't up \& shew'd his Face;        
  In spite of Darknes, it was DAY.       
    It was \textit{Thy} day, \textit{Sweet}! \& did rise        
  Not from the \textit{East}, but from thine \textit{Eyes}.        

{\Forma \strophae {}
 \1 It was \textit{Thy} day, Sweet, \&c.
}       

    \3 \textit{Winter} chidde aloud; \& sent        
  The angry North to wage his warres.       
    The North forgott his feirce Intent;        
  And left perfumes in stead of scarres.        
    By those sweet eyes persuasive powrs        
  Where he mean't frost, he scatter'd flowrs.        
       
{\Forma \strophae {}
 \1 By those sweet eyes, \&c.  
}       

    \4 We saw thee in thy baulmy Nest,        
  Young dawn of our eternall \textit{Day}!        
    We saw thine eyes break from their \textit{East}        
  And chase the trembling shades away.        
    We saw thee; \& we blest the sight,       
  We saw thee by thine own sweet light.        
       
    \2 Poor \textit{World} (said I) what wilt thou doe        
  To entertain this starry \textit{Stranger}?        
    Is this the best thou canst bestow?        
  A cold, and not too cleanly, manger?       
    Contend ye powres of heav'n \& earth        
  To fitt a bed for this huge birthe.        
       
{\Forma \strophae {}
 \1 Contend ye powers, \&c.
}      

    \3 Proud world, said I; cease your contest,        
  And let the \textit{Mighty} \textit{Babe} alone.       
    The Phoenix builds the Phoenix' nest.        
  Lov's architecture is his own.        
    The \textit{Babe} whose birth embraves this morn,        
  Made his own bed e're he was born.        
       
{\Forma \strophae {}
 \1 The \textit{Babe} whose, \&c.
}       

    \2 I saw the curl'd drops, soft \& slow,        
  Come hovering o're the place's head;        
    Offring their whitest sheets of snow        
  To furnish the fair \textit{Infant}'s bed:        
    Forbear, said I; be not too bold.       
  Your fleece is white, But t'is too cold.        
       
{\Forma \strophae {} 
 \1 Forbear, sayd I, \&c.
}       

    \3 I saw the obsequious \textit{Seraphins}        
  Their rosy fleece of fire bestow,        
    For well they now can spare their wings,       
  Since \textit{Heaven} it self lyes here below.        
    Well done, said I: but are you sure        
  Your down so warm, will passe for pure?        
       
{\Forma \strophae {} 
 \1 Well done sayd I, \&c. 
}       

    \2 No no, your \textit{King}'s not yet to seeke       
  Where to repose his Royall \textit{Head},        
    See see, how soon his new-bloom'd \textit{Cheek}        
  Twixt's mother's brests is gone to bed.        
    Sweet choise, said we! no way but so        
  Not to ly cold, yet sleep in snow.       
       
{\Forma \strophae {}
 \1 Sweet choise, said we, \&c. 
}       

    \4 We saw thee in thy baulmy nest,        
  Bright dawn of our eternall Day!        
    We saw thine eyes break from their \textit{East}        
  And chase the trembling shades away.       
    We saw thee: \& we blest the sight.        
  We saw thee, by thine own sweet light.        
       
{\Forma \strophae {}
 \1 We saw thee, \&c.
}

       
  \5       Wellcome, all \textit{Wonders} in one sight!        
  Eternity shutt in a span.       
    Sommer in Winter. Day in Night.        
  Heaven in earth, \& \textit{God} in \textit{Man}.        
    Great little one! whose all-embracing birth        
  Lifts earth to heaven, stoopes heav'n to earth.        
       
    \textit{Wellcome}. Though nor to gold nor silk.       
  To more then Caesar's birth right is;        
    Two sister-seas of Virgin-Milk,        
  With many a rarely-temper'd kisse        
    That breathes at once both \textit{Maid} \& \textit{Mother},        
  Warmes in the one, cooles in the other.       
       
    \textit{Wellcome}, though not to those gay flyes        
  Guilded ith' Beames of earthly kings;        
    Slippery soules in smiling eyes;        
  But to poor Shepherds, home-spun things:        
    Whose Wealth 's their flock; whose witt, to be       
  Well read in their simplicity.        
       
    Yet when young April's husband showrs        
  Shall blesse the fruitfull Maia's bed,        
    We'l bring the First-born of her flowrs        
  To kisse thy \textit{Feet} \& crown thy \textit{Head}.       
    To thee, dread Lamb! whose love must keep        
  The shepheards, more then they the sheep.        
       
    To \textit{Thee} meek Majesty! soft \textit{King}        
  Of simple \textit{Graces} \& sweet \textit{Loves}.        
    Each of us his lamb will bring       
  Each his pair of sylver Doves;        
    Till burnt at last in fire of Thy fair eyes,        
  Our selves become our own best \textit{Sacrifice}.        
\endDrama
\end{poem}      

\poet{Andrew Marvell\indexes{Marvell, Andrew}}{1621\textendash1678}
      
\begin{poem}{A Dialogue between\\ The Resolved Soul, and Created Pleasure}

Courage my Soul, now learn to wield  \auto
The weight of thine immortal Shield.
Close on thy Head thy Helmet bright.
Ballance thy Sword against the Fight.
See where an Army, strong as fair,
With silken Banners spreads the air.
Now, if thou bee'st that thing Divine,
In this day's Combat let it shine:
And shew that Nature wants an Art
To conquer one resolved Heart.

\Drama
\Facies \personae {\textsc{#1}}
\Locus            {\centre}
\SpatiumSupra     {1ex plus .25ex minus .25ex\penalty -100}
\SpatiumInfra     {.33ex plus .11ex\penalty 10000}
\persona*[1]{pleasure}
\persona*[2]{soul}
\persona*[3]{chorus}

\1 Welcome the Creations Guest,
Lord of Earth, and Heavens Heir.
Lay aside that Warlike Crest,
And of Nature's banquet share:
Where the Souls of fruits and flow'rs
Stand prepar'd to heighten yours.

\2 I sup above, and cannot stay
To bait so long upon the way.

\1 On these downy Pillows lye,
Whose soft Plumes will thither fly:
On these Roses strow'd so plain
Lest one Leaf thy Side should strain.

\2 My gentler Rest is on a Thought,
Conscious of doing what I ought.

\1 If thou bee'st with Perfumes pleas'd,
Such as  oft the Gods appeas'd,
Thou in fragrant Clouds shalt show
Like another God below.

\2 A Soul that knowes not to presume
Is Heaven's and its own perfume.

\1 Every thing does seem to vie
Which should first attract thine Eye:
But since none deserves that grace,
In this Crystal view thy face.

\2 When the Creator's skill is priz'd,
The rest is all but Earth disguis'd.

\1 Heark how Musick then prepares
For thy Stay these charming Aires ;
Which the posting Winds recall,
And suspend the Rivers Fall.

\2 Had I but any time to lose,
On this I would it all dispose.
Cease Tempter. None can chain a mind
Whom this sweet Chordage cannot bind.

{\itshape
\3  Earth  cannot shew so brave a Sight
As when a single Soul does fence
The Batteries of alluring Sense,
And Heaven views it with delight.
Then persevere: for still new Charges sound:
And if thou overcom'st thou shalt be crown'd.
}

\1 All this fair, and cost, and sweet,
Which scatteringly doth shine,
Shall within one Beauty meet,
And she be only thine.

\2 If things of Sight such Heavens be,
What Heavens are those we cannot see?

\1 Where so e're thy Foot shall go
The minted Gold shall lie;
Till thou purchase all below,
And want new Worlds to buy.

\2 Wer't not a price who 'ld value Gold?
And that's worth nought that can be sold.

\1 Wilt thou all the Glory have
That War or Peace commend?
Half the World shall be thy Slave
The other half thy Friend.

\2 What Friends, if to my self untrue?
What Slaves, unless I captive you?

\1 Thou shalt know each hidden Cause;
And see the future Time:
Try what depth the Centre draws;
And then to Heaven climb.

\2 None thither mounts by the degree
Of Knowledge, but Humility.

{\itshape
\3 Triumph, triumph, victorious Soul;
The World has not one Pleasure more:
The rest does lie beyond the pole,
And is thine everlasting Store.
}
\end{poem}

\begin{poem}{On a Drop of Dew}
\Forma\strophae{0010011001012{-1}0110200222002*{10}0*3 \LastButOne 0}%
See how the Orient Dew,
Shed from the Bosom of the Morn \auto
Into the blowing Roses,
Yet careless of its Mansion new;
For the clear Region where 'twas born  
Round in its self incloses:
And in its little Globes Extent,
Frames as it can its native Element.
How it the purple flow'r does slight,
Scarce touching where it lyes,
But gazing back upon the Skies,
Shines with a mournful Light;
Like its own Tear,
Because so long divided from the Sphear.
Restless it roules and unsecure,
Trembling lest it grow impure:
Till the warm Sun pitty it's Pain,
And to the Skies exhale it back again.
So the Soul, that Drop, that Ray
Of the clear Fountain of Eternal Day,
Could it within the humane flow'r be seen,
Remembring still its former height,
Shuns the sweat leaves and blossoms green;
And, recollecting its own Light,
Does, in its pure and circling thoughts, express
The greater Heaven in an Heaven less.
In how coy a Figure wound,
Every way it turns away:
So the World excluding round,
Yet receiving in the Day.
Dark beneath, but bright above:
Here disdaining, there in Love.
How loose and easie hence to go:
How girt and ready to ascend.
Moving but on a point below,
It all about does upwards bend.
Such did the Manna's sacred Dew destil;
White, and intire, though congeal'd and chill.
Congeal'd on Earth: but does, dissolving, run
Into the Glories of th' Almighty Sun.
\endDrama
\end{poem}
\begin{poem}{The Coronet}   
\Forma\strophae{0110101001101111\LastButOne0}%
      
  When for the Thorns with which I long, too long, \auto       
          With many a piercing wound,        
          My Saviours head have crown'd,        
  I seek with Garlands to redress that Wrong:        
      Through every Garden, every Mead,       
  I gather flow'rs (my fruits are only flow'rs)        
      Dismantling all the fragrant Towers        
  That once adorn'd my Shepherdesses head.        
  And now when I have summ'd up all my store,        
      Thinking (so I my self deceive)       
      So rich a Chaplet thence to weave        
  As never yet the king of Glory wore:        
      Alas I find the Serpent old        
      That, twining in his speckled breast,        
      About the flow'rs disguis'd does fold,       
      With wreaths of Fame and Interest.        
\Forma\strophae{}%
  Ah, foolish Man, that would'st debase with them,        
  And mortal Glory, Heavens Diadem!        
  But thou who only could'st the Serpent tame,        
  Either his slipp'ry knots at once untie,       
  And disintangle all his winding Snare:        
  Or shatter too with him my curious frame:        
  And let these wither, so that he may die,        
  Though set with Skill and chosen out with Care.        
  That they, while Thou on both their Spoils dost tread,       
  May crown thy Feet, that could not crown thy Head.        
\end{poem}
\endDrama
\clearpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\newpagestyle {IndexAuthorsPage} {

  \sethead   [\oldstylenums{\thepage}]
             [\textsc{index of authors}]
             []
             {}
             {\textsc{index of authors}}
             {\oldstylenums{\thepage}}
}

\TypesetIndex [a] {% The index of authors

\pagestyle {IndexAuthorsPage}

\Facies       \tituli  {\thispagestyle {empty}\MakeUppercase{#1}}
\SpatiumInfra          {2\leading plus 1\leading minus .5\leading}
\Caput                 {\section}

\titulus {Index of Authors}

\Facies \aIndexEntryText {#1}
\Modus                   {\leaders}

\Facies \aIndexEntryPage {\oldstylenums{#1}}

}{\clearpage}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newpagestyle {IndexTitlesPage} {

  \sethead   [\oldstylenums{\thepage}]
             [\textsc{index of titles}]
             []
             {}
             {\textsc{index of titles}}
             {\oldstylenums{\thepage}}
}

\TypesetIndex {% The index of titles

\pagestyle {IndexTitlesPage}

\Facies       \tituli  {\thispagestyle {empty}\MakeUppercase{#1}}
\SpatiumInfra          {2\leading plus 1\leading minus .5\leading}
\Caput                 {\section}

\titulus {Index of Titles}

\Facies \tIndexEntryText {#1}
\Modus                   {\leaders}

\Facies \tIndexEntryPage {\oldstylenums{#1}}

\Facies \lIndexEntryText {#1}
\Modus                   {\leaders}

\Facies \lIndexEntryPage {\oldstylenums{#1}}

}{\clearpage}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newpagestyle {IndexFirstPage} {

  \sethead   [\oldstylenums{\thepage}]
             [\textsc{index of first lines}]
             []
             {}
             {\textsc{index of first lines}}
             {\oldstylenums{\thepage}}
}

\TypesetIndex [l] {% The index of of first lines

\pagestyle {IndexFirstPage}

\Facies       \tituli  {\thispagestyle {empty}\MakeUppercase{#1}}
\SpatiumInfra          {2\leading plus 1\leading minus .5\leading}
\Caput                 {\section}

\titulus {Index of First Lines}

\Facies \lIndexEntryText {#1}

\Facies \lIndexEntryPage {\oldstylenums{#1}}

}{\clearpage}


\end{document}

