\documentclass[DIV=9, parskip=half, pagesize=auto]{scrartcl}

\usepackage{fixltx2e}
\usepackage{etex}
\usepackage{xspace}
\usepackage{lmodern}
\usepackage[T1]{fontenc}
\usepackage{textcomp}
\usepackage{array}
\usepackage{microtype}
\usepackage{hyperref}

\newcommand*{\mail}[1]{\href{mailto:#1}{\texttt{#1}}}
\newcommand*{\pkg}[1]{\textsf{#1}}
\newcommand*{\cls}[1]{\textsf{#1}}
\newcommand*{\cs}[1]{\texttt{\textbackslash#1}}
\makeatletter
\newcommand*{\cmd}[1]{\cs{\expandafter\@gobble\string#1}}
\makeatother
\newcommand*{\env}[1]{\texttt{#1}}
\newcommand*{\opt}[1]{\texttt{#1}}
\newcommand*{\myllap}[1]{%
  \quitvmode
  \setbox0=\hbox{#1}%
  \kern-\wd0
  #1%
}

\deffootnote{1em}{0em}{\textsuperscript{\thefootnotemark}\,}
\addtokomafont{title}{\rmfamily}

\title{The \pkg{listing} package\thanks{This manual corresponds to \pkg{listing}~V1.2, dated~1999/05/25.}}
\subtitle{\LaTeX\ package for typesetting listings.}
\author{%
  Matthew Hebley\thanks{Matthew is no longer available for contact, email or otherwise}%
  \and Volker Kuhlmann\thanks{University of Canterbury\newline EEE Dept\newline Christchurch, New Zealand\newline E-Mail: \mail{v.kuhlmann@elec.canterbury.ac.nz}}%
}
\date{1999/05/25}


\begin{document}

\maketitle

\begin{quote}
  \footnotesize
  This program can be redistributed and/or modified under the terms
  of the \LaTeX\ Project Public License, distributed from CTAN
  archives as \href{http://www.ctan.org/pub/tex-archive/macros/latex/base/lppl.txt}{\texttt{macros/latex/base/lppl.txt}}; either
  version~1 of the License, or (at your option) any later version.
\end{quote}


\minisec{History:}

\begin{tabular}{@{}ll>{\raggedright\arraybackslash}p{10cm}@{}}
  1.2 & 25 May 1999 & Re-published under the LPPL\@. No other change.                                           \\
  1.1 & 07 Feb 96   & Added setting of \cmd{\thelisting} with the \cls{book} and \cls{report} document classes. \\
  1.0 & 19 Aug 94   & Created.
\end{tabular}


\minisec{Load with:}

\verb+\usepackage{listing}+\\
or\\
\verb+\usepackage[bigcaptions]{listing}+


\minisec{Package option:}

\begin{labeling}{\opt{bigcaptions}}
\item[\opt{bigcaptions}] If present, captions appear as for sections
  (or chapters when using the ``\cls{book}'' or ``\cls{report}'' 
  classes).
  If not present, captions appear as for captions
  in the \env{figure} or \env{table} environment.
\end{labeling}


\minisec{Available control sequences:}

\begin{labeling}{\cmd{\listlistingname}}
\item[\cmd{\begin\{listing\}}] A new environment for listings is provided.
  Unlike \env{figure} and \env{table}, listing is \textls[40]{NOT}
  a floating environment.
\item[\cmd{\caption}] As usual. (only inside listing!)
\item[\cmd{\label}] As usual. (only inside listing; after \cmd{\caption}!)
\item[\cmd{\ref}] As usual.
\item[\cmd{\listoflistings}] As for \cmd{\listoftables}, \cmd{\listoffigures}.
\item[\cmd{\listingname}] Produces the ``Listing'' text for the caption.
\item[\cmd{\listlistingname}] Produces the ``List of Listings'' text.
\end{labeling}

A \texttt{*.lol} file is produced similar to \texttt{*.lot}, \texttt{*.lof} files.

Does not change \cmd{\@makecaption} (unless option ``bigcaptions'' is used),
so packages that change the layout of \cmd{\caption} still work.
\pkg{listing.sty} works together with \pkg{hangcaption.sty}.

\begin{quote}
  \myllap{``}Documentation is like sex.
  If it is good, it is \textls[50]{REALLY} good, 
  and if it is bad, it is better than nothing!''
  
  \hfill[Gabe Helou]
\end{quote}

\end{document}
