%\documentclass[12pt,english]{hletter}
\documentclass[12pt,english,signit]{hletter}
%\documentclass[11pt,special,french,draft]{hletter}

\usepackage{mergeh}

%% to make a private letter use a "from address":
%\address{57, Main Street \\ Upper Throgmortendale\\ \textbf{York  YO23 3QS} \\
%    England\\[1ex]  tel: +44 1671 425 6382\\ mobile: +44 1762 4253 4123}

\signature{A. Nother\\ Head of Batology Dept.}

%% for the \closingtwo (at end) use:
%\signature{Dr.~C. Featherstonehaugh & Dr.~A. Beauchamp \\ CEO & CIO}

\date{Lausanne, le 15 septembre 2008}

\begin{document}
\sign[10mm]{signat}  %  \sign may be here or in the option file

\reference{Impressions of the journey from Vevey to Lausanne.}

\begin{merge}{testmerge.dat}
Cully, midway between Vevey and Lausanne, is a little town of 1,100
inhabitants. It can look back very far into the past, for there was a
time when the lounger on the beach of Cully would have seen in the bay
a lake village, an amphibious camp made up of thatched roofs and huts
on a submerged field of piles. Later on the idler would have seen Roman
legionaries and Roman merchants halting by the road, since Cully seems
to have been a posting station or wayside caravanserai of some
significance.

Cully, from its earliest days, has been devoted to the making of wine.
It has probably made wine ever since wine-making was known in Europe.
It makes wine still and makes it well and, indeed, does nothing else.
Grapes and the winepress are the symbols of its being, the subject of
its thoughts and the mark of its ambition. A bunch of grapes on a
branch constitutes the ancient arms of Cully. They are grapes, it may
be noticed, of such size and quality that they would seem to proclaim
to the world, \emph{these are the grapes of Cully.}  Someone, inspired
by a sense of the fitness of things, has surmised that there was once a
temple to Bacchus at Cully, but unfortunately the suggestion is
unfounded.

 \closing{Yours Sincerely,}

\vfill
\cc{All Smiths in London\\ Mademoiselle S. Curchod}

\encl{Tourist guide to Switzerland.\\ Plan of Cully.}

\end{merge}

\end{document}
